---
title: "Distance Project"
author: "Nigel K Anderson"
format: html
editor: visual
---

```{r setup, include=FALSE}
library(tidyverse)

bat_theme <- theme_classic(base_size = 15, base_family = "Arial") +
  theme(
    plot.background    = element_rect(fill = "white", color = NA),
    panel.background   = element_rect(fill = "white", color = NA),
    panel.grid.major.y = element_line(color = "#EBEBEB", linewidth = 0.4),
    panel.grid.major.x = element_blank(),
    axis.line          = element_line(color = "#444444", linewidth = 0.5),
    axis.ticks         = element_line(color = "#444444"),
    axis.text          = element_text(color = "#333333", size = 16),
    axis.title         = element_text(color = "#222222", size = 18),
    plot.title         = element_text(color = "#111111", size = 14,
                                      face = "bold", margin = margin(b = 3)),
    plot.subtitle      = element_text(color = "#666666", size = 10,
                                      margin = margin(b = 8)),
    legend.background  = element_rect(fill = "white", color = NA),
    legend.key         = element_rect(fill = "white", color = NA),
    legend.text        = element_text(color = "#333333", size = 10),
    legend.title       = element_text(color = "#444444", size = 9, face = "bold"),
    strip.background   = element_rect(fill = "#F2F2F2", color = NA),
    strip.text         = element_text(color = "#222222", face = "bold", size = 10),
    plot.caption       = element_text(color = "#999999", size = 8,
                                      hjust = 0, margin = margin(t = 8)),
    plot.margin        = margin(12, 14, 10, 12)
  )

col_red    <- "#7A1C2E"   
col_white  <- "#C9A84C"   
col_navy   <- "#1B2A4A"   
col_purple <- "#4A1259"   
col_ivory  <- "#F5F0E3"   

intensity_pal <- c("10"  = col_ivory,
                   "30"  = col_white,
                   "50"  = col_purple,
                   "70"  = col_red,
                   "100" = col_navy)
```

## Data Pipeline

```{r import, warning=FALSE, message=TRUE}
source("R/01_import_sonobat.R")
```

```{r clean, warning=FALSE, message=FALSE}
source("R/02_clean_data.R")
data <- clean_data(data_raw)
```

```{r habitat, warning=FALSE, message=FALSE}
source("R/03_extract_habitat.R")

data_env <- add_habitat(
  data,
  forest_raster = nlcd_forest,
  buffer = 50
)

data_env <- add_moonlight(data_env)
```

```{r treatment, warning=FALSE, message=FALSE}
source("R/07_add_light_treatment.R")

data_env <- data_env %>%
  left_join(
    lighting %>% select(date, color, intensity),
    by = "date"
  ) %>%
  mutate(
    color     = factor(color),
    site      = factor(site),
    intensity = factor(intensity)
  )

```

```{r distance, warning=FALSE, message=FALSE}
library(sf)

ref_point <- st_as_sf(
  data.frame(lon = -110.645134, lat = 43.904879),
  coords = c("lon", "lat"),
  crs = 4326
)

sites_sf <- st_as_sf(sites, coords = c("lon", "lat"), crs = 4326)
sites_sf <- st_transform(sites_sf, 32612)
ref_point <- st_transform(ref_point, 32612)

sites_sf$dist_m <- as.numeric(st_distance(sites_sf, ref_point))

# Pull only dist_m 
site_dists <- as.data.frame(sites_sf) %>% dplyr::select(site, dist_m)

data_env <- data_env %>%
  left_join(site_dists, by = "site") %>%
  mutate(
    dist_km = dist_m / 1000,
    dist_cat = case_when(
      dist_km < 0.52 ~ "Close",
      dist_km < 1    ~ "Medium",
      dist_km < 2.1  ~ "Further",
      TRUE           ~ "Far"
    ),
    jd_c = jd - mean(jd, na.rm = TRUE)   # centered Julian day
  )

data_env <- data_env %>% filter(!is.na(dist_km))

data_env %>% select(site, dist_km) %>% distinct() %>% arrange(dist_km)

# Reclassify out-of-region misidentifications onto the correct regional species,
# INCLUDING inside compound labels: Labo (Eastern red bat) -> Mylu and Myse
# (N. long-eared) -> Myev. Token-wise replacement; duplicate candidates
# that result are collapsed while preserving vote order, so e.g. "Mylu/Labo" ->
# "Mylu" (becomes a single-species call) and "Lano/Myse/Mylu" -> "Lano/Myev/Mylu".
.fix_aliases <- function(x) {
  vapply(stringr::str_split(x, "/"), function(toks) {
    toks <- dplyr::recode(toks, Labo = "Mylu", Myse = "Myev")
    paste(toks[!duplicated(toks)], collapse = "/")
  }, character(1))
}
data_env <- data_env %>% mutate(species = .fix_aliases(species))

# Merging can put two rows on the same species at the same site/date/recording;
# sum detections so counts stay additive and re-derive mean confidence.
.recode_key <- c("site", "year", "date", "jd", "jd2", "species",
                 "datetime", "time_sec")
data_env <- data_env %>%
  group_by(across(all_of(.recode_key))) %>%
  summarise(detections          = sum(detections),
            weighted_detections = sum(weighted_detections),
            across(-c(detections, weighted_detections, mean_confidence), first),
            .groups = "drop") %>%
  mutate(mean_confidence = weighted_detections / detections)

# Focal single-species set for the per-species and community models.
focal_spp <- c("Epfu", "Laci", "Lano", "Myca", "Myci",
               "Mylu", "Myvo", "Myyu", "Myev")

# Acoustic guilds by call frequency. 
lowfreq_spp  <- c("Epfu", "Lano", "Laci")                  # low frequency
medfreq_spp  <- c("Myev")                                  # medium frequency (Myev only)
highfreq_spp <- c("Mylu", "Myvo", "Myci", "Myca", "Myyu")  # high frequency
assign_cluster <- function(x) {
  vapply(stringr::str_split(x, "/"), function(toks) {
    if (all(toks %in% lowfreq_spp))       "LowFreq"
    else if (all(toks %in% medfreq_spp))  "Myev"
    else if (all(toks %in% highfreq_spp)) "HighFreq"
    else NA_character_
  }, character(1))
}

# Overall stream = focal single species + compounds that fall within ONE guild.
# Cross-guild compounds (e.g. Mylu/Myev, Anpa/Epfu) and non-focal singles are
# dropped, so the kept compounds match the guild definitions and the overall
# total equals the sum of the three guilds. Report what gets dropped.
.dropped <- data_env %>%
  filter(is.na(assign_cluster(species))) %>%
  mutate(kind = ifelse(str_detect(species, "/"),
                       "cross-guild compound", "non-focal single")) %>%
  group_by(kind) %>%
  summarise(labels = n_distinct(species), detections = sum(detections),
            .groups = "drop")
message("Dropped (not within a single guild): ",
        paste(sprintf("%s = %d detections (%d labels)",
                      .dropped$kind, .dropped$detections, .dropped$labels),
              collapse = " | "))

data_env <- data_env %>% filter(!is.na(assign_cluster(species)))

message("Overall data rows: ", nrow(data_env),
        "  |  single-species detections: ",
        sum(data_env$detections[!str_detect(data_env$species, "/")]),
        "  |  within-guild compound labels: ",
        n_distinct(data_env$species[str_detect(data_env$species, "/")]),
        " (", sum(data_env$detections[str_detect(data_env$species, "/")]),
        " detections)")
```

```{r sky_brightness, warning=FALSE, message=TRUE}
source("R/09_add_sky_brightness.R")
```

```{r bat_traits, warning=FALSE, message=FALSE}
library(readxl)

bat_traits <- read_excel("data/bat_traits.xlsx") %>%
  rename(species = Species)

data_env <- data_env %>%
  left_join(bat_traits, by = "species")

# Single-species subset for the per-species and community models. Compound
# (multi-species) labels are excluded here but remain in `data_env` for the
# overall model and figures.
data_env_focal <- data_env %>% filter(species %in% focal_spp)

# Per-site-night totals: bat activity = sum of detections across every species /
# compound label on a site-night. The site-night is the unit the light treatment
# is applied to.
# Every covariate is constant within a site-night, so nothing is lost aggregating.
site_night <- data_env %>%
  group_by(site, date) %>%
  summarise(
    total_det       = sum(detections, na.rm = TRUE),
    color           = first(color),
    intensity       = first(intensity),
    dist_km         = first(dist_km),
    mean_phase      = first(mean_phase),
    pct_nonforest   = first(pct_nonforest),
    brightness_dark = first(brightness_dark),
    jd              = first(jd),
    jd_c            = first(jd_c),
    .groups = "drop"
  )
```

## Species Detection Counts

```{r species_counts, warning=FALSE, message=FALSE, fig.width=8, fig.height=5}
# Total detections per species (compound multi-species labels pooled into one bar)
spp_counts <- data_env %>%
  mutate(grp = ifelse(str_detect(species, "/"), "(compound)", species)) %>%
  group_by(grp) %>%
  summarise(detections = sum(detections), .groups = "drop") %>%
  mutate(grp = fct_reorder(grp, detections))

fig_spp_det <- ggplot(spp_counts, aes(detections, grp)) +
  geom_col(fill = col_navy, width = 0.75) +
  geom_text(aes(label = scales::comma(detections)),
            hjust = -0.1, size = 3.5, color = "#333333") +
  scale_x_continuous(expand = expansion(mult = c(0, 0.18)),
                     labels = scales::comma) +
  labs(title = "Total detections per species",
       subtitle = "SonoBat v30.1, Prob >= 0.9; compound (multi-species) labels pooled",
       x = "Total detections", y = NULL) +
  bat_theme

fig_spp_det

# Prevalence: how many of the sampled site-nights each focal species appears on
spp_prev <- data_env_focal %>%
  distinct(species, site, date) %>%
  count(species, name = "site_nights") %>%
  mutate(species = fct_reorder(species, site_nights))

fig_spp_prev <- ggplot(spp_prev, aes(site_nights, species)) +
  geom_col(fill = col_purple, width = 0.75) +
  geom_text(aes(label = site_nights), hjust = -0.15, size = 3.5, color = "#333333") +
  scale_x_continuous(expand = expansion(mult = c(0, 0.12))) +
  labs(title = "Site-nights each species was detected",
       subtitle = paste0("Out of ", n_distinct(paste(data_env$site, data_env$date)),
                         " sampled site-nights (focal species)"),
       x = "Site-nights with at least one detection", y = NULL) +
  bat_theme

fig_spp_prev

dir.create("output/figures", showWarnings = FALSE, recursive = TRUE)
ggsave("output/figures/species_counts.png", fig_spp_det,
       width = 8, height = 5, dpi = 200, bg = "white")
ggsave("output/figures/species_prevalence.png", fig_spp_prev,
       width = 8, height = 5, dpi = 200, bg = "white")
```

## Models

### Overall activity model (total detections per night)

```{r simple_model, warning=FALSE, message=FALSE}
library(glmmTMB)

# Overall activity model at the site-night level (one row per site x night).
# Total detections per night ~ treatment + covariates, with a site random effect.
# No ziformula: nightly totals have no structural zeros.
simple_model1 <- glmmTMB(
  total_det ~
    color * intensity +
    dist_km * intensity +
    mean_phase +
    pct_nonforest +
    brightness_dark +
    jd_c + I(jd_c^2) +
    (1 | site),
  data   = site_night,
  family = nbinom2()
)

summary(simple_model1)
```

## Guild-cluster models

Rather than forcing ambiguous compound calls onto a single species, group every detection into the acoustic guild by call frequency: **low** (Epfu / Lano / Laci), **medium** (Myev), and **high** (Mylu / Myvo / Myci / Myca / Myyu). The medium-frequency guild is a single species, so it is labelled **Myev**. A detection joins a guild only if *every* candidate species in its label belongs to that guild. `data_env` holds only within-guild labels and the three guilds sum to the overall total. This models activity at the level SonoBat can actually resolve and it separates guilds that respond to distance differently (low-frequency activity concentrates near the lit parking lot, while the medium- and high-frequency guilds are more active farther out), which a pooled total would cancel out.

```{r cluster_setup, warning=FALSE, message=FALSE}
# `assign_cluster` and the guild sets (lowfreq_spp / medfreq_spp / highfreq_spp)
# are defined in the data pipeline above, where data_env was already restricted
# to within-guild labels -- so every row here assigns to a guild.
cl_det <- data_env %>%
  mutate(cluster = assign_cluster(species)) %>%
  group_by(site, date, cluster) %>%
  summarise(detections = sum(detections), .groups = "drop")

# zero-filled: every sampled site-night x guild, carrying site-night covariates
cluster_night <- site_night %>%
  select(site, date, color, intensity, dist_km, mean_phase,
         pct_nonforest, brightness_dark, jd_c) %>%
  tidyr::crossing(cluster = c("LowFreq", "Myev", "HighFreq")) %>%
  left_join(cl_det, by = c("site", "date", "cluster")) %>%
  mutate(detections = tidyr::replace_na(detections, 0),
         site = factor(site),
         cluster = factor(cluster, levels = c("LowFreq", "Myev", "HighFreq")))

# guild x colour x intensity, plus a guild-specific distance slope
cluster_glmm <- glmmTMB(
  detections ~ color * intensity + cluster * dist_km + color +
    mean_phase + pct_nonforest + brightness_dark + jd_c + I(jd_c^2) +
    (1 | site),
  data      = cluster_night,
  ziformula = ~1,
  family    = nbinom2()
)

summary(cluster_glmm)

# guild-specific distance slopes (log scale)
emmeans::emtrends(cluster_glmm, ~ cluster, var = "dist_km")
```

### Guild responses to intensity × color and to distance

```{r cluster_figs, warning=FALSE, message=FALSE, fig.width=13, fig.height=5}
library(emmeans)

# (1) intensity x color by guild
emm_cl <- as.data.frame(
  emmeans(cluster_glmm, ~ color * intensity | cluster, type = "response"))
names(emm_cl)[names(emm_cl) %in% c("response", "rate", "emmean")] <- "est"
pd <- position_dodge(0.35)
fig_cluster_ic <- ggplot(
    emm_cl, aes(factor(intensity), est, color = color, group = color)) +
  geom_line(linewidth = 0.9, position = pd) +
  geom_point(size = 2.5, position = pd) +
  geom_errorbar(aes(ymin = asymp.LCL, ymax = asymp.UCL),
                width = 0.2, linewidth = 0.7, position = pd) +
  scale_color_manual(values = c(R = col_red, W = col_white),
                     labels = c(R = "Red", W = "White"), name = "Light color") +
  facet_wrap(~ cluster, scales = "free_y") +
  labs(title = "Guild response to intensity × color",
       subtitle = "Predicted detections/night ± 95% CI",
       x = "Light intensity (%)", y = "Predicted detections / night") +
  bat_theme

# (2) distance response by guild (other covariates at their means, at 50% red)
dgrid <- expand.grid(
  dist_km   = seq(min(cluster_night$dist_km), max(cluster_night$dist_km), length = 60),
  cluster   = factor(c("LowFreq", "Myev", "HighFreq"), levels = levels(cluster_night$cluster)),
  color     = factor("R",  levels = levels(cluster_night$color)),
  intensity = factor("50", levels = levels(cluster_night$intensity)),
  jd_c = 0,
  mean_phase      = mean(cluster_night$mean_phase, na.rm = TRUE),
  pct_nonforest   = mean(cluster_night$pct_nonforest, na.rm = TRUE),
  brightness_dark = mean(cluster_night$brightness_dark, na.rm = TRUE),
  site = NA)
pr <- predict(cluster_glmm, newdata = dgrid, se.fit = TRUE, re.form = NA, type = "link")
dgrid$fit <- exp(pr$fit)
dgrid$lo  <- exp(pr$fit - 1.96 * pr$se.fit)
dgrid$hi  <- exp(pr$fit + 1.96 * pr$se.fit)
fig_cluster_dist <- ggplot(dgrid, aes(dist_km, fit, color = cluster, fill = cluster)) +
  geom_ribbon(aes(ymin = lo, ymax = hi), alpha = 0.15, color = NA) +
  geom_line(linewidth = 1.1) +
  scale_color_manual(values = c(LowFreq = col_navy, Myev = col_purple, HighFreq = col_red), name = NULL) +
  scale_fill_manual(values = c(LowFreq = col_navy, Myev = col_purple, HighFreq = col_red), guide = "none") +
  scale_y_log10() +   # guilds span ~2 orders of magnitude; log keeps all slopes visible
  labs(title = "Guild response to distance",
       subtitle = "Predicted detections/night ± 95% CI (at 50% red, other covariates at means); log scale",
       x = "Distance from Colter Bay parking lot (km)",
       y = "Predicted detections / night (log)") +
  bat_theme

fig_cluster_ic
fig_cluster_dist

ggsave("output/figures/cluster_intensity_color.png", fig_cluster_ic,
       width = 11, height = 5, dpi = 200, bg = "white")
ggsave("output/figures/cluster_distance.png", fig_cluster_dist,
       width = 8, height = 5, dpi = 200, bg = "white")
```

## Species models

### Community-average species model (hierarchical)

Total detections are \~68% `Mylu`, so a pooled total mostly reflects one species. This model instead keeps one row per site-night **× focal species** (with zeros filled in for species not detected that night) and enters species as a random effect, so the fixed effects estimate the **community-average** light response rather than a `Mylu`-weighted total. The zero-filled cells (\~39%) make the zero-inflated NB appropriate.

```{r species_glmm, warning=FALSE, message=FALSE}
library(glmmTMB)

# focal single-species detections summed to the site-night level
spp_det <- data_env_focal %>%
  group_by(site, date, species) %>%
  summarise(detections = sum(detections), .groups = "drop")

# zero-filled grid: every sampled site-night x focal species (0 where undetected),
# carrying the site-night covariates from `site_night`
spp_night <- site_night %>%
  select(site, date, color, intensity, dist_km, mean_phase,
         pct_nonforest, brightness_dark, jd_c) %>%
  tidyr::crossing(species = focal_spp) %>%
  left_join(spp_det, by = c("site", "date", "species")) %>%
  mutate(detections = tidyr::replace_na(detections, 0),
         site    = factor(site),
         species = factor(species))

# Random intercepts for site and species. (A per-species colour slope,
# (1 + color | species), was singular here, so intercepts only.)
species_glmm <- glmmTMB(
  detections ~
    color * intensity +
    dist_km +
    mean_phase +
    pct_nonforest +
    brightness_dark +
    jd_c + I(jd_c^2) +
    (1 | site) + (1 | species),
  data      = spp_night,
  ziformula = ~1,
  family    = nbinom2()
)

summary(species_glmm)
```

### Data-integrity sanity checks

Assert the core invariants of the bat data pipeline before interpreting any model. These fail loudly (`stopifnot`) if the data being processed is not what the analysis assumes — e.g. an out-of-region alias slipped through, a guild set changed, a zero-fill grid is the wrong size, or a join dropped/duplicated site-nights.

```{r sanity_checks_bats, warning=FALSE, message=TRUE}
.n_sn  <- nrow(dplyr::distinct(data_env, site, date))
.toks  <- unlist(strsplit(unique(as.character(data_env$species)), "/"))

stopifnot(
  # focal set & guild partition
  "focal_spp should be 9 species"                 = length(focal_spp) == 9,
  "Myth must not be in focal_spp"                 = !("Myth" %in% focal_spp),
  "guild sets must be disjoint"                    = length(intersect(lowfreq_spp, c(medfreq_spp, highfreq_spp))) == 0 &&
                                                       length(intersect(medfreq_spp, highfreq_spp)) == 0,
  "guild sets must partition focal_spp"            = setequal(c(lowfreq_spp, medfreq_spp, highfreq_spp), focal_spp),
  "medium-frequency guild is Myev only"           = identical(sort(medfreq_spp), "Myev"),

  # alias reclassification removed out-of-region tokens (Labo->Mylu, Myse->Myev)
  "Labo should be reclassified away"              = !("Labo" %in% .toks),
  "Myse should be reclassified away"              = !("Myse" %in% .toks),

  # every retained label maps to exactly one guild (cluster filter worked)
  "every data_env label maps to a guild"          = all(!is.na(assign_cluster(data_env$species))),

  # distance & treatment sanity
  "distances are non-negative"                    = all(data_env$dist_km >= 0, na.rm = TRUE),
  "distances within plausible range (<= 8 km)"    = all(data_env$dist_km <= 8, na.rm = TRUE),
  "intensity has the 5 expected levels"           = setequal(levels(data_env$intensity), c("10","30","50","70","100")),
  "colour is R/W"                                 = setequal(levels(data_env$color), c("R","W")),

  # site_night: one row per site-night, totals preserved, no missing covariates/treatment
  "site_night has one row per site-night"         = nrow(site_night) == .n_sn,
  "site_night preserves total detections"         = sum(site_night$total_det) == sum(data_env$detections),
  "site_night covariates complete"                = all(!is.na(dplyr::select(site_night, dist_km, mean_phase, pct_nonforest, brightness_dark))),
  "site_night treatment complete"                 = all(!is.na(site_night$color)) && all(!is.na(site_night$intensity)),

  # cluster_night: zero-filled site-night x 3 guilds, totals preserved
  "cluster_night is site-night x 3 guilds"        = nrow(cluster_night) == .n_sn * 3,
  "cluster_night has no NA detections"            = all(!is.na(cluster_night$detections)),
  "guild totals equal the overall total"          = sum(cluster_night$detections) == sum(data_env$detections),

  # spp_night: zero-filled site-night x focal species, focal totals preserved
  "spp_night is site-night x focal species"       = nrow(spp_night) == .n_sn * length(focal_spp),
  "spp_night has no NA detections"                = all(!is.na(spp_night$detections)),
  "spp_night preserves focal-species totals"      = sum(spp_night$detections) == sum(data_env_focal$detections)
)

message("Bat sanity checks passed: ", .n_sn, " site-nights across ",
        dplyr::n_distinct(data_env$site), " sites; ", length(focal_spp),
        " focal species; 3 guilds; distance ",
        paste(round(range(data_env$dist_km, na.rm = TRUE), 2), collapse = "-"), " km.")
```

### Model-based intensity × color effects

Figure-3-style effect plots, but from the fitted models: predicted response by light intensity and colour with 95% CIs (estimated marginal means, other covariates held at their means, averaged over the random effects).

```{r model_effect_figs, warning=FALSE, message=FALSE, fig.width=8, fig.height=9}
library(emmeans)
library(patchwork)

# predicted response by intensity x color (+/- 95% CI), in the Fig 3 style
emm_intensity_color <- function(model, ttl, sub, ylab) {
  pd <- position_dodge(0.3)
  e  <- as.data.frame(emmeans(model, ~ color * intensity, type = "response"))
  names(e)[names(e) %in% c("response", "rate", "emmean")] <- "est"
  ggplot(e, aes(factor(intensity), est, color = color, group = color)) +
    geom_line(linewidth = 1, position = pd) +
    geom_point(size = 3, position = pd) +
    geom_errorbar(aes(ymin = asymp.LCL, ymax = asymp.UCL),
                  width = 0.15, linewidth = 0.8, position = pd) +
    scale_color_manual(values = c(R = col_red, W = col_white),
                       labels = c(R = "Red", W = "White"), name = "Light color") +
    labs(title = ttl, subtitle = sub, x = "Light intensity (%)", y = ylab) +
    bat_theme
}

fig_eff_species <- emm_intensity_color(
  species_glmm, "Community-average species response",
  "Model-based EMM +/- 95% CI (species_glmm)",
  "Predicted detections / species / night")

fig_eff_total <- emm_intensity_color(
  simple_model1, "Overall nightly activity",
  "Model-based EMM +/- 95% CI (simple_model1)",
  "Predicted total detections / night")

fig_eff_species / fig_eff_total

ggsave("output/figures/model_intensity_color.png",
       fig_eff_species / fig_eff_total,
       width = 8, height = 9, dpi = 200, bg = "white")
```

### Per-species intensity × color effects

Each species gets its own **intensity** response, with a **common** colour effect. Likelihood-ratio tests show species differ strongly in their intensity response (species × intensity, p \< 1e-16) but *not* in their colour response (species × colour, p = 0.43), and the full species × colour × intensity three-way is not worth its parameters (AIC and BIC both prefer dropping it). So we use `species * intensity + colour * intensity` — supported, better-fitting, and interpretable without a three-way. The red-vs-white gap is therefore consistent across species; the panels differ in their intensity shapes.

```{r model_effect_byspecies, warning=FALSE, message=FALSE, fig.width=13, fig.height=6.5}
library(emmeans)

# species-specific intensity response + a common colour x intensity effect
# (the 3-way species:color:intensity was significant but AIC/BIC-worse, and
# species:color was n.s.); same zero-filled grid as species_glmm
species_glmm_int <- glmmTMB(
  detections ~ species * intensity + color * intensity +
    dist_km + mean_phase + pct_nonforest + brightness_dark + jd_c + I(jd_c^2) +
    (1 | site),
  data      = spp_night,
  ziformula = ~1,
  family    = nbinom2()
)

emm_sp <- as.data.frame(
  emmeans(species_glmm_int, ~ color * intensity | species, type = "response"))
names(emm_sp)[names(emm_sp) %in% c("response", "rate", "emmean")] <- "est"

pd <- position_dodge(0.4)
fig_eff_byspecies <- ggplot(
    emm_sp, aes(factor(intensity), est, color = color, group = color)) +
  geom_line(linewidth = 0.8, position = pd) +
  geom_point(size = 2, position = pd) +
  geom_errorbar(aes(ymin = asymp.LCL, ymax = asymp.UCL),
                width = 0.2, linewidth = 0.6, position = pd) +
  scale_color_manual(values = c(R = col_red, W = col_white),
                     labels = c(R = "Red", W = "White"), name = "Light color") +
  facet_wrap(~ species, scales = "free_y", nrow = 2) +
  labs(title = "Per-species intensity effect, with a common colour effect (model-based)",
       subtitle = "Predicted detections/night ± 95% CI; species × intensity + colour × intensity (no 3-way); y-axes free",
       x = "Light intensity (%)", y = "Predicted detections / night") +
  bat_theme

fig_eff_byspecies

ggsave("output/figures/model_intensity_color_byspecies.png", fig_eff_byspecies,
       width = 13, height = 6.5, dpi = 200, bg = "white")
```

### Per-species distance response

Model-predicted detections across the distance gradient for each focal species, from the same zero-filled `spp_night` grid, now with a `species × dist_km` interaction (a common colour and intensity effect). **Caveat:** unlike the moth distance figures, the bat distance signal is fragile — it hinges on the three far (\~7 km) sites and GRTE15 in particular (see the robustness section), so read these as a near-vs-far *sort* (low-frequency bats near, Myotis far), not firm per-species slopes.

```{r species_distance, warning=FALSE, message=FALSE, fig.width=12, fig.height=6}
bat_spp_dist_glmm <- glmmTMB(
  detections ~ species * dist_km + color + intensity + mean_phase +
    pct_nonforest + brightness_dark + jd_c + I(jd_c^2) + (1 | site),
  data = spp_night, ziformula = ~1, family = nbinom2())

cat("=== Per-species distance slope (log detections / km) ===\n")
print(as.data.frame(emtrends(bat_spp_dist_glmm, ~ species, var = "dist_km")) %>%
        transmute(species, slope = round(dist_km.trend, 2), SE = round(SE, 2)))

# per-species distance-slope p-values with multiple-comparison correction (9 species)
.sl <- as.data.frame(emtrends(bat_spp_dist_glmm, ~ species, var = "dist_km"))
.p  <- 2 * pnorm(-abs(.sl$dist_km.trend / .sl$SE))
knitr::kable(tibble::tibble(Species = as.character(.sl$species),
                            Slope = round(.sl$dist_km.trend, 3),
                            p_raw = round(.p, 3),
                            p_Holm = round(p.adjust(.p, "holm"), 3),
                            p_FDR = round(p.adjust(.p, "BH"), 3)) %>% dplyr::arrange(Slope),
             caption = "Bat per-species distance slopes: raw and multiple-comparison-corrected p-values.")

dist_grid <- seq(min(spp_night$dist_km), max(spp_night$dist_km), length.out = 80)
bat_spp_pred <- as.data.frame(
  emmeans(bat_spp_dist_glmm, ~ species | dist_km,
          at = list(dist_km = dist_grid), type = "response")) %>%
  rename(lower = asymp.LCL, upper = asymp.UCL) %>%
  mutate(species = factor(species, levels = focal_spp))

fig_bat_species_distance <- ggplot(bat_spp_pred, aes(dist_km, response)) +
  geom_ribbon(aes(ymin = lower, ymax = upper), fill = col_navy, alpha = 0.15) +
  geom_line(color = col_navy, linewidth = 1) +
  facet_wrap(~ species, scales = "free_y", nrow = 2) +
  labs(title = "Predicted Bat Detections Across Distance, by Species",
       subtitle = "Model-estimated marginal means ± 95% CI (negative-binomial GLMM)",
       x = "Distance from parking lot (km)",
       y = "Predicted detections per site-night") +
  bat_theme
fig_bat_species_distance

ggsave("output/figures/bat_species_distance_panels.png", fig_bat_species_distance,
       width = 12, height = 6, dpi = 300, bg = "white")

bat_spp_pal <- setNames(
  c("#1B2A4A","#7A1C2E","#C9A84C","#4A1259","#2E7D5B","#B5651D",
    "#3B6EA5","#9C3848","#6B8E23","#7A5195"), focal_spp)
fig_bat_species_distance_combined <- ggplot(bat_spp_pred,
    aes(dist_km, response, color = species, fill = species)) +
  geom_ribbon(aes(ymin = lower, ymax = upper), alpha = 0.08, color = NA) +
  geom_line(linewidth = 1) + scale_y_log10() +
  scale_color_manual(values = bat_spp_pal, name = "Species") +
  scale_fill_manual(values = bat_spp_pal, name = "Species") +
  labs(title = "Predicted Bat Detections Across Distance, by Species",
       subtitle = "Model-estimated marginal means (log scale)",
       x = "Distance from parking lot (km)",
       y = "Predicted detections per site-night (log scale)") +
  bat_theme
fig_bat_species_distance_combined

ggsave("output/figures/bat_species_distance_combined.png",
       fig_bat_species_distance_combined,
       width = 9, height = 6, dpi = 300, bg = "white")
```

Robustness: refit the same per-species distance model with the three far (\~7 km) sites removed (GRTE13/14/15). The distance range collapses to 0–2 km and most slopes **flip sign** — the Myotis that rose toward far sites now decline with distance within the near cluster — confirming that the bat distance gradient is carried by the far sites rather than being a smooth trend.

```{r species_distance_nofar, warning=FALSE, message=FALSE, fig.width=12, fig.height=6}
spp_night_nofar <- spp_night %>%
  filter(!site %in% c("GRTE13","GRTE14","GRTE15")) %>% droplevels()

bat_spp_dist_glmm_nofar <- glmmTMB(
  detections ~ species * dist_km + color + intensity + mean_phase +
    pct_nonforest + brightness_dark + jd_c + I(jd_c^2) + (1 | site),
  data = spp_night_nofar, ziformula = ~1, family = nbinom2())

cat("=== Per-species distance slope, FAR SITES REMOVED (log detections / km) ===\n")
print(as.data.frame(emtrends(bat_spp_dist_glmm_nofar, ~ species, var = "dist_km")) %>%
        transmute(species, slope = round(dist_km.trend, 2), SE = round(SE, 2)))

# per-species distance-slope p-values with multiple-comparison correction (9 species)
.sln <- as.data.frame(emtrends(bat_spp_dist_glmm_nofar, ~ species, var = "dist_km"))
.pn  <- 2 * pnorm(-abs(.sln$dist_km.trend / .sln$SE))
knitr::kable(tibble::tibble(Species = as.character(.sln$species),
                            Slope = round(.sln$dist_km.trend, 3),
                            p_raw = round(.pn, 3),
                            p_Holm = round(p.adjust(.pn, "holm"), 3),
                            p_FDR = round(p.adjust(.pn, "BH"), 3)) %>% dplyr::arrange(Slope),
             caption = "Bat per-species distance slopes, far sites removed: raw and multiple-comparison-corrected p-values.")

dist_grid_nofar <- seq(min(spp_night_nofar$dist_km), max(spp_night_nofar$dist_km),
                       length.out = 80)
bat_spp_pred_nofar <- as.data.frame(
  emmeans(bat_spp_dist_glmm_nofar, ~ species | dist_km,
          at = list(dist_km = dist_grid_nofar), type = "response")) %>%
  rename(lower = asymp.LCL, upper = asymp.UCL) %>%
  mutate(species = factor(species, levels = focal_spp))

fig_bat_species_distance_nofar <- ggplot(bat_spp_pred_nofar, aes(dist_km, response)) +
  geom_ribbon(aes(ymin = lower, ymax = upper), fill = col_navy, alpha = 0.15) +
  geom_line(color = col_navy, linewidth = 1) +
  facet_wrap(~ species, scales = "free_y", nrow = 2) +
  labs(title = "Predicted Bat Detections Across Distance, by Species",
       subtitle = "Far sites (GRTE13/14/15) removed — marginal means ± 95% CI",
       x = "Distance from parking lot (km)",
       y = "Predicted detections per site-night") +
  bat_theme
fig_bat_species_distance_nofar

ggsave("output/figures/bat_species_distance_panels_nofar.png",
       fig_bat_species_distance_nofar,
       width = 12, height = 6, dpi = 300, bg = "white")

fig_bat_species_distance_combined_nofar <- ggplot(bat_spp_pred_nofar,
    aes(dist_km, response, color = species, fill = species)) +
  geom_ribbon(aes(ymin = lower, ymax = upper), alpha = 0.08, color = NA) +
  geom_line(linewidth = 1) + scale_y_log10() +
  scale_color_manual(values = bat_spp_pal, name = "Species") +
  scale_fill_manual(values = bat_spp_pal, name = "Species") +
  labs(title = "Predicted Bat Detections Across Distance, by Species",
       subtitle = "Far sites (GRTE13/14/15) removed — marginal means (log scale)",
       x = "Distance from parking lot (km)",
       y = "Predicted detections per site-night (log scale)") +
  bat_theme
fig_bat_species_distance_combined_nofar

ggsave("output/figures/bat_species_distance_combined_nofar.png",
       fig_bat_species_distance_combined_nofar,
       width = 9, height = 6, dpi = 300, bg = "white")
```

### Overall bat activity: intensity/colour GAM

GAM counterpart to the moth intensity/colour GAM, on total nightly bat activity (`site_night$total_det`): intensity as a smooth with a separate curve per colour, site as a random effect. Only 5 distinct intensity values, so the smooth basis is capped at `k = 5`.

```{r bat_gam_intensity_color, warning=FALSE, message=FALSE, fig.width=8, fig.height=5.5}
library(mgcv)
bat_gam_dat <- site_night %>%
  mutate(intensity_num = as.numeric(as.character(intensity))) %>%
  filter(!is.na(intensity_num), !is.na(color))

bat_gam <- gam(
  total_det ~ color + s(intensity_num, by = color, k = 5) +
    s(dist_km, k = 5) + s(jd_c, k = 5) +
    mean_phase + pct_nonforest + brightness_dark + s(site, bs = "re"),
  data = bat_gam_dat, family = nb(), method = "REML")
cat("=== Overall bat intensity/colour GAM ===\n"); print(summary(bat_gam))

bat_gam_shared <- gam(
  total_det ~ color + s(intensity_num, k = 5) + s(dist_km, k = 5) + s(jd_c, k = 5) +
    mean_phase + pct_nonforest + brightness_dark + s(site, bs = "re"),
  data = bat_gam_dat, family = nb(), method = "REML")
cat("\nAIC (separate-by-colour vs shared intensity smooth):\n")
print(AIC(bat_gam_shared, bat_gam))

bat_pgrid <- expand_grid(intensity_num = seq(10, 100, by = 1),
                         color = factor(c("R","W"), levels = levels(bat_gam_dat$color))) %>%
  mutate(dist_km = mean(bat_gam_dat$dist_km), jd_c = 0,
         mean_phase = mean(bat_gam_dat$mean_phase, na.rm = TRUE),
         pct_nonforest = mean(bat_gam_dat$pct_nonforest, na.rm = TRUE),
         brightness_dark = mean(bat_gam_dat$brightness_dark, na.rm = TRUE),
         site = bat_gam_dat$site[1])
bpr <- predict(bat_gam, newdata = bat_pgrid, type = "link", se.fit = TRUE, exclude = "s(site)")
bilink <- family(bat_gam)$linkinv
bat_pgrid <- bat_pgrid %>% mutate(fit = bilink(bpr$fit),
                                  lo = bilink(bpr$fit - 1.96*bpr$se.fit),
                                  hi = bilink(bpr$fit + 1.96*bpr$se.fit))
bat_obs <- bat_gam_dat %>% group_by(color, intensity_num) %>%
  summarise(mean_det = mean(total_det), .groups = "drop")

fig_bat_gam <- ggplot(bat_pgrid, aes(intensity_num, fit, color = color, fill = color)) +
  geom_ribbon(aes(ymin = lo, ymax = hi), alpha = 0.15, color = NA) +
  geom_line(linewidth = 1.1) +
  geom_point(data = bat_obs, aes(intensity_num, mean_det, color = color),
             inherit.aes = FALSE, size = 2.4, alpha = 0.8) +
  scale_color_manual(values = c(R = col_red, W = col_white),
                     labels = c(R = "Red", W = "White"), name = "Light color") +
  scale_fill_manual(values = c(R = col_red, W = col_white),
                    labels = c(R = "Red", W = "White"), name = "Light color") +
  labs(title = "Bat Activity vs. Light Intensity by Color (GAM)",
       subtitle = "Negative-binomial GAM, smooth of intensity per color ± 95% CI; points = observed site-night means",
       x = "Light intensity (%)", y = "Predicted detections per site-night") +
  bat_theme + theme(legend.position = "top")
fig_bat_gam

ggsave("output/figures/bat_gam_intensity_color.png", fig_bat_gam,
       width = 8, height = 5.5, dpi = 300, bg = "white")
```

## Community Analysis

```{r community, warning=FALSE, message=FALSE}
source("R/10_community_analysis.R")
community <- run_community_analysis(data_env_focal, bat_theme, col_red, col_white)
```

```{r community_results, warning=FALSE, message=FALSE}
# NMDS stress
cat("NMDS stress:", round(community$nmds$stress, 3), "\n")

# PERMANOVA
community$perm

# Homogeneity of dispersion (by color)
community$disp_test

# dbRDA — global and marginal tests
community$db_global
community$db_margin

# mvabund — multivariate GLM
community$mv_anova
```

```{r community_figs, warning=FALSE, message=FALSE}
community$fig_c1   # dbRDA ordination biplot
community$fig_c2   # Species activity heatmap (color × intensity)
community$fig_c3   # Species activity heatmap (distance)
```

## Moth (Insect) Data Import

```{r insect_import, warning=FALSE, message=FALSE}
source("R/13_import_insects.R")
insects <- import_insects("data/grte_distance_insectID.xlsx")
```

```{r insect_env, warning=FALSE, message=FALSE}
insect_events <- insects$total %>%
  left_join(sites, by = "site") %>%
  left_join(site_dists, by = "site") %>%
  mutate(
    dist_km  = dist_m / 1000,
    dist_cat = case_when(
      dist_km < 0.52 ~ "Close",
      dist_km < 1    ~ "Medium",
      dist_km < 2.1  ~ "Further",
      TRUE           ~ "Far"
    ),
    jd = lubridate::yday(date)
  ) %>%
  filter(!is.na(dist_km))

insect_events <- add_habitat(insect_events, forest_raster = nlcd_forest, buffer = 50,
                              cache_path = "data/insect_data_out.rds")

# Entries are dated but not exactly timestamped; noon avoids any timezone date-rollover
insect_events$datetime <- as.POSIXct(paste(insect_events$date, "12:00:00"), tz = "America/Denver")
insect_events <- add_moonlight(insect_events)

sky <- read_csv("data/sky_brightness_measurements.csv", show_col_types = FALSE) %>%
  mutate(site = toupper(site))
insect_events <- insect_events %>%
  left_join(sky %>% select(site, brightness_dark), by = "site")

insect_total_env <- insect_events %>%
  mutate(
    jd_c      = jd - mean(jd, na.rm = TRUE),
    color     = factor(color),
    site      = factor(site),
    intensity = factor(intensity)
  )

message("Moth sampling events: ", nrow(insect_total_env),
        "  |  Sites: ", n_distinct(insect_total_env$site),
        "  |  Date range: ", paste(range(insect_total_env$date), collapse = " to "))
```

### Data-integrity sanity checks (moths)

Assert the moth data is what the models assume before fitting. The moth unit is a sampling **event** (site $\times$ date $\times$ colour $\times$ intensity); note one date has two events at a site, so there are slightly more events than unique site-dates.

```{r sanity_checks_moths, warning=FALSE, message=TRUE}
stopifnot(
  # one row per sampling event; counts are non-negative integers
  "insect_total_env is one row per event"   = nrow(insect_total_env) ==
                                                 dplyr::n_distinct(paste(insect_total_env$site, insect_total_env$date,
                                                                         insect_total_env$color, insect_total_env$intensity)),
  "moth detections are non-negative integers" = all(insect_total_env$detections >= 0) &&
                                                 all(insect_total_env$detections == floor(insect_total_env$detections)),

  # covariates present and in range
  "moth distances present and in range"     = all(!is.na(insect_total_env$dist_km)) &&
                                                 all(insect_total_env$dist_km >= 0 & insect_total_env$dist_km <= 8),
  "moth site-night covariates complete"     = all(!is.na(insect_total_env$mean_phase)) &&
                                                 all(!is.na(insect_total_env$pct_nonforest)) &&
                                                 all(!is.na(insect_total_env$brightness_dark)),

  # treatment factor levels
  "moth intensity has 5 expected levels"    = setequal(as.character(unique(insect_total_env$intensity)), c("10","30","50","70","100")),
  "moth colour is R/W"                      = setequal(as.character(unique(insect_total_env$color)), c("R","W")),

  # taxonomic detection frames are consistent with the specimen table
  "family counts == specimens with family" = sum(insects$family$detections) == sum(!is.na(insects$specimens$Family)),
  "family detections positive integers"    = all(insects$family$detections >= 1) &&
                                                 all(insects$family$detections == floor(insects$family$detections)),
  "specimen table is non-empty"            = nrow(insects$specimens) > 0
)

message("Moth sanity checks passed: ", nrow(insect_total_env), " events across ",
        dplyr::n_distinct(insect_total_env$site), " sites; ",
        dplyr::n_distinct(insects$family$Family), " families; ",
        nrow(insects$specimens), " specimens.")
```

## Moth General Activity Model

```{r insect_general_model, warning=FALSE, message=FALSE}
source("R/15_insect_models.R")
insect_general <- run_insect_general_model(insect_total_env, bat_theme, intensity_pal)

summary(insect_general$model)
```

```{r insect_general_posthoc, warning=FALSE, message=FALSE}
cat("=== Color effect within each intensity level ===\n")
insect_general$em_color_intensity
pairs(insect_general$em_color_intensity, adjust = "holm")

cat("\n=== Distance slope by intensity level ===\n")
insect_general$em_dist_trend
```

```{r insect_general_figs, warning=FALSE, message=FALSE}
insect_general$fig_color_intensity
insect_general$fig_dist_intensity
```

```{r insect_gam_intensity_color, warning=FALSE, message=FALSE, fig.width=8, fig.height=5.5}
# GAM alternative to the factor model: intensity as a smooth, a separate curve
# per colour, so the intensity response can bend and differ by colour. Only 5
# distinct intensity values, so the smooth basis is capped at k = 5. Site enters
# as a random effect; the other covariates are controls.
library(mgcv)
gam_dat <- insect_total_env %>%
  mutate(intensity_num = as.numeric(as.character(intensity))) %>%
  filter(!is.na(intensity_num), !is.na(color))

moth_gam <- gam(
  detections ~ color + s(intensity_num, by = color, k = 5) +
    s(dist_km, k = 5) + s(jd_c, k = 5) +
    mean_phase + pct_nonforest + brightness_dark + s(site, bs = "re"),
  data = gam_dat, family = nb(), method = "REML")
cat("=== Moth intensity/colour GAM ===\n"); print(summary(moth_gam))

# does the intensity smooth differ by colour? (vs. a single shared smooth)
moth_gam_shared <- gam(
  detections ~ color + s(intensity_num, k = 5) + s(dist_km, k = 5) + s(jd_c, k = 5) +
    mean_phase + pct_nonforest + brightness_dark + s(site, bs = "re"),
  data = gam_dat, family = nb(), method = "REML")
cat("\nAIC (separate-by-colour vs shared intensity smooth):\n")
print(AIC(moth_gam_shared, moth_gam))

pgrid <- expand_grid(intensity_num = seq(10, 100, by = 1),
                     color = factor(c("R","W"), levels = levels(gam_dat$color))) %>%
  mutate(dist_km = mean(gam_dat$dist_km), jd_c = 0,
         mean_phase = mean(gam_dat$mean_phase, na.rm = TRUE),
         pct_nonforest = mean(gam_dat$pct_nonforest, na.rm = TRUE),
         brightness_dark = mean(gam_dat$brightness_dark, na.rm = TRUE),
         site = gam_dat$site[1])
pr <- predict(moth_gam, newdata = pgrid, type = "link", se.fit = TRUE, exclude = "s(site)")
ilink <- family(moth_gam)$linkinv
pgrid <- pgrid %>% mutate(fit = ilink(pr$fit),
                          lo = ilink(pr$fit - 1.96 * pr$se.fit),
                          hi = ilink(pr$fit + 1.96 * pr$se.fit))
obs <- gam_dat %>% group_by(color, intensity_num) %>%
  summarise(mean_det = mean(detections), .groups = "drop")

fig_moth_gam <- ggplot(pgrid, aes(intensity_num, fit, color = color, fill = color)) +
  geom_ribbon(aes(ymin = lo, ymax = hi), alpha = 0.15, color = NA) +
  geom_line(linewidth = 1.1) +
  geom_point(data = obs, aes(intensity_num, mean_det, color = color),
             inherit.aes = FALSE, size = 2.4, alpha = 0.8) +
  scale_color_manual(values = c(R = col_red, W = col_white),
                     labels = c(R = "Red", W = "White"), name = "Light color") +
  scale_fill_manual(values = c(R = col_red, W = col_white),
                    labels = c(R = "Red", W = "White"), name = "Light color") +
  labs(title = "Moth Activity vs. Light Intensity by Color (GAM)",
       subtitle = "Negative-binomial GAM, smooth of intensity per color ± 95% CI; points = observed site-night means",
       x = "Light intensity (%)", y = "Predicted detections per site-night") +
  bat_theme + theme(legend.position = "top")
fig_moth_gam

ggsave("output/figures/moth_gam_intensity_color.png", fig_moth_gam,
       width = 8, height = 5.5, dpi = 300, bg = "white")
```

## Moth Community Analysis (Family)

```{r insect_family_env, warning=FALSE, message=FALSE}
event_covariates <- insect_events %>%
  select(site, date, color, intensity, dist_km, dist_cat, jd,
         mean_phase, mean_moonlight, max_moonlight,
         pct_forest, pct_nonforest, brightness_dark)

insect_family_env <- insects$family %>%
  left_join(event_covariates, by = c("site", "date", "color", "intensity")) %>%
  filter(!is.na(dist_km)) %>%
  rename(species = Family) %>%
  mutate(color = factor(color), site = factor(site), intensity = factor(intensity))

message("Moth family-level rows: ", nrow(insect_family_env),
        "  |  Families: ", paste(sort(unique(insect_family_env$species)), collapse = ", "))
```

```{r insect_genus_species_distance, warning=FALSE, message=FALSE, fig.width=10, fig.height=6.5}
# Same across-distance predictions at subfamily, genus, and species resolution.
# Restrict to the 9 best-sampled taxa (>= 30 site-nights) at each level; for
# genus, drop labels that are really subfamily/family placeholders (-inae/-idae).
source("R/33_moth_taxonomy.R")   # add_species_bin()
sp_tax <- add_species_bin(insects$specimens)

make_taxon_distance <- function(level, out_tag, pretty,
                                exclude_rx = NULL, n_top = 9, min_sitenights = 30,
                                drop_sites = NULL) {
  sfx <- if (is.null(drop_sites)) "" else "_nofar"
  sub_note <- if (is.null(drop_sites))
    "Model-estimated marginal means ± 95% CI (negative-binomial GLMM)" else
    "Far sites (GRTE13/14/15) removed — marginal means ± 95% CI"
  sub_note_log <- if (is.null(drop_sites))
    "Model-estimated marginal means (log scale)" else
    "Far sites (GRTE13/14/15) removed — marginal means (log scale)"
  det <- sp_tax %>% filter(!is.na(.data[[level]])) %>% rename(taxon = all_of(level))
  if (!is.null(exclude_rx)) det <- det %>% filter(!str_detect(taxon, exclude_rx))
  keep <- det %>% group_by(taxon) %>%
    summarise(d = n(), sn = n_distinct(paste(site, date)), .groups = "drop") %>%
    filter(sn >= min_sitenights) %>% slice_max(d, n = n_top) %>% arrange(desc(d)) %>%
    pull(taxon)

  det_kept <- det %>% filter(taxon %in% keep) %>%
    group_by(site, date, color, intensity, taxon) %>%
    summarise(detections = n(), .groups = "drop")

  grid <- event_covariates %>%
    mutate(jd_c = jd - mean(jd, na.rm = TRUE)) %>%
    tidyr::crossing(taxon = keep) %>%
    left_join(det_kept, by = c("site","date","color","intensity","taxon")) %>%
    mutate(detections = tidyr::replace_na(detections, 0),
           taxon = factor(taxon, levels = keep), site = factor(site),
           color = factor(color),
           intensity = factor(intensity, levels = c("10","30","50","70","100")))
  if (!is.null(drop_sites))
    grid <- grid %>% filter(!site %in% drop_sites) %>% droplevels()

  fit <- function(extra) glmmTMB(
    as.formula(paste("detections ~ taxon * dist_km + color + intensity + mean_phase +",
                     extra, "jd_c + I(jd_c^2) + (1 | site)")),
    data = grid, ziformula = ~1, family = nbinom2())
  stable <- function(m) isTRUE(m$sdr$pdHess) &&
    !any(is.na(suppressWarnings(sqrt(diag(vcov(m)$cond)))))
  m <- fit("pct_nonforest + brightness_dark +")
  if (!stable(m)) m <- fit("brightness_dark +")
  if (!stable(m)) m <- fit("")
  cat("\n===", pretty, "distance slopes (log detections / km) ===\n")
  print(as.data.frame(emtrends(m, ~ taxon, var = "dist_km")) %>%
          transmute(taxon, slope = round(dist_km.trend, 2), SE = round(SE, 2)))

  dg <- seq(min(grid$dist_km), max(grid$dist_km), length.out = 80)
  pred <- as.data.frame(emmeans(m, ~ taxon | dist_km, at = list(dist_km = dg),
                                type = "response")) %>%
    rename(lower = asymp.LCL, upper = asymp.UCL) %>%
    mutate(taxon = factor(taxon, levels = keep))

  fig_fac <- ggplot(pred, aes(dist_km, response)) +
    geom_ribbon(aes(ymin = lower, ymax = upper), fill = col_navy, alpha = 0.15) +
    geom_line(color = col_navy, linewidth = 1) +
    facet_wrap(~ taxon, scales = "free_y") +
    labs(title = paste0("Predicted Moth Detections Across Distance, by ", pretty),
         subtitle = sub_note,
         x = "Distance from parking lot (km)",
         y = "Predicted detections per site-night") +
    bat_theme + theme(strip.text = element_text(face = "bold.italic", size = 11))
  ggsave(sprintf("output/figures/moth_%s_distance_panels%s.png", out_tag, sfx), fig_fac,
         width = 10, height = 6.5, dpi = 300, bg = "white")

  pal <- setNames(c("#1B2A4A","#7A1C2E","#C9A84C","#4A1259","#2E7D5B","#B5651D",
                    "#3B6EA5","#9C3848","#6B8E23")[seq_along(keep)], keep)
  fig_all <- ggplot(pred, aes(dist_km, response, color = taxon, fill = taxon)) +
    geom_ribbon(aes(ymin = lower, ymax = upper), alpha = 0.10, color = NA) +
    geom_line(linewidth = 1) + scale_y_log10() +
    scale_color_manual(values = pal, name = pretty) +
    scale_fill_manual(values = pal, name = pretty) +
    labs(title = paste0("Predicted Moth Detections Across Distance, by ", pretty),
         subtitle = sub_note_log,
         x = "Distance from parking lot (km)",
         y = "Predicted detections per site-night (log scale)") +
    bat_theme + theme(legend.text = element_text(face = "italic"))
  ggsave(sprintf("output/figures/moth_%s_distance_combined%s.png", out_tag, sfx), fig_all,
         width = 9, height = 6, dpi = 300, bg = "white")
  print(fig_fac)
}

# full data
make_taxon_distance("Subfamily",   "subfamily", "Subfamily")
make_taxon_distance("Genus",       "genus",   "Genus",   exclude_rx = "(inae|idae)$")
make_taxon_distance("Species_bin", "species", "Species")

# robustness: three far (~7 km) sites removed
far3 <- c("GRTE13","GRTE14","GRTE15")
make_taxon_distance("Subfamily",   "subfamily", "Subfamily", drop_sites = far3)
make_taxon_distance("Genus",       "genus",   "Genus",   exclude_rx = "(inae|idae)$",
                    drop_sites = far3)
make_taxon_distance("Species_bin", "species", "Species", drop_sites = far3)
```

### Significant-taxa distance hierarchy

Synthesises the per-taxon distance results into one taxonomic tree. At each level (family → subfamily → genus → species) we test every taxon with ≥30 site-nights for a distance slope (joint `taxon × dist_km` NB-GLMM), keep those with *p* < 0.05, and draw each as a miniature of its model-predicted detections-vs-distance curve (±95% CI, auto-scaled within its box). Nodes are coloured by direction (more active away from vs. toward the lot) and linked to their nearest significant ancestor. *p* is uncorrected across the many taxa, so single-star nodes should be read cautiously.

```{r moth_sig_hierarchy, warning=FALSE, message=FALSE, fig.width=14, fig.height=8}
# taxonomic lineage lookups (modal parent per taxon)
.tax <- sp_tax %>% filter(!is.na(Family)) %>% count(Family, Subfamily, Genus, Species_bin, name = "nn")
.first_by <- function(df, key, cols) df %>% group_by(across(all_of(key))) %>%
  slice_max(nn, n = 1, with_ties = FALSE) %>% ungroup() %>% select(all_of(c(key, cols)))
.sub2 <- .first_by(.tax %>% filter(!is.na(Subfamily)) %>% count(Subfamily, Family, wt = nn, name = "nn"), "Subfamily", "Family")
.gen2 <- .first_by(.tax %>% filter(!is.na(Genus)) %>% count(Genus, Subfamily, Family, wt = nn, name = "nn"), "Genus", c("Subfamily", "Family"))
.sp2  <- .first_by(.tax %>% filter(!is.na(Species_bin)) %>% count(Species_bin, Genus, Subfamily, Family, wt = nn, name = "nn"), "Species_bin", c("Genus", "Subfamily", "Family"))
.stbl <- function(m) !is.null(m) && isTRUE(m$sdr$pdHess) && !any(is.na(suppressWarnings(sqrt(diag(vcov(m)$cond)))))
.pstar <- function(p) ifelse(is.na(p), "", ifelse(p < .001, "***", ifelse(p < .01, "**", ifelse(p < .05, "*", ""))))
INT_LV <- c("10", "30", "50", "70", "100")

# per-level distance model -> per-taxon slope/p and predicted distance curve
hier_level <- function(level, drop_sites = NULL, exclude_rx = NULL, min_sn = 30) {
  det <- sp_tax %>% filter(!is.na(.data[[level]])) %>% rename(taxon = all_of(level))
  if (!is.null(exclude_rx)) det <- det %>% filter(!str_detect(taxon, exclude_rx))
  keep <- det %>% group_by(taxon) %>% summarise(sn = n_distinct(paste(site, date)), .groups = "drop") %>%
    filter(sn >= min_sn) %>% pull(taxon)
  if (length(keep) < 1) return(NULL)
  det_kept <- det %>% filter(taxon %in% keep) %>%
    group_by(site, date, color, intensity, taxon) %>% summarise(detections = n(), .groups = "drop")
  grid <- event_covariates %>% mutate(jd_c = jd - mean(jd, na.rm = TRUE)) %>%
    tidyr::crossing(taxon = keep) %>%
    left_join(det_kept, by = c("site", "date", "color", "intensity", "taxon")) %>%
    mutate(detections = tidyr::replace_na(detections, 0), taxon = factor(taxon, levels = keep),
           site = factor(site), color = factor(color), intensity = factor(intensity, levels = INT_LV))
  if (!is.null(drop_sites)) grid <- grid %>% filter(!site %in% drop_sites) %>% droplevels()
  cand <- c("mean_phase + pct_nonforest + brightness_dark + jd_c + I(jd_c^2)",
            "mean_phase + brightness_dark + jd_c + I(jd_c^2)", "mean_phase + jd_c + I(jd_c^2)",
            "mean_phase + jd_c", "jd_c")
  m <- NULL
  for (rhs in cand) {
    mi <- tryCatch(suppressWarnings(glmmTMB(
      as.formula(paste("detections ~ taxon * dist_km + color + intensity +", rhs, "+ (1 | site)")),
      data = grid, ziformula = ~1, family = nbinom2())), error = function(e) NULL)
    if (is.null(m) && !is.null(mi)) m <- mi
    if (.stbl(mi)) { m <- mi; break }
  }
  if (is.null(m)) return(NULL)
  s <- as.data.frame(emtrends(m, ~ taxon, var = "dist_km"))
  sumry <- tibble(taxon = as.character(s$taxon), est = s$dist_km.trend,
                  p = 2 * pnorm(-abs(s$dist_km.trend / s$SE)))
  dg <- seq(min(grid$dist_km), max(grid$dist_km), length.out = 40)
  cv <- as.data.frame(emmeans(m, ~ taxon | dist_km, at = list(dist_km = dg), type = "response")) %>%
    transmute(taxon = as.character(taxon), dist_km, response, lower = asymp.LCL, upper = asymp.UCL)
  list(sumry = sumry, curves = cv)
}

hier_nodes <- function(drop_sites = NULL) {
  lv <- list(Family = hier_level("Family", drop_sites), Subfamily = hier_level("Subfamily", drop_sites),
             Genus = hier_level("Genus", drop_sites, exclude_rx = "(inae|idae)$"),
             Species_bin = hier_level("Species_bin", drop_sites))
  yy <- c(Family = 4, Subfamily = 3, Genus = 2, Species_bin = 1)
  nd <- bind_rows(lapply(names(lv), function(l) { x <- lv[[l]]; if (is.null(x)) return(NULL)
    x$sumry %>% filter(!is.na(p), p < 0.05) %>% mutate(lev = l, y = yy[l]) }))
  curves <- bind_rows(lapply(names(lv), function(l) { x <- lv[[l]]; if (is.null(x)) return(NULL)
    x$curves %>% mutate(lev = l) }))
  if (nrow(nd) == 0) return(list(nd = nd, curves = curves))
  nd$Family <- NA; nd$Subfamily <- NA; nd$Genus <- NA; nd$Species_bin <- NA
  for (i in seq_len(nrow(nd))) { t <- nd$taxon[i]; l <- nd$lev[i]
    if (l == "Family") nd$Family[i] <- t
    if (l == "Subfamily") { nd$Subfamily[i] <- t; r <- .sub2[.sub2$Subfamily == t, ]; if (nrow(r)) nd$Family[i] <- r$Family[1] }
    if (l == "Genus") { nd$Genus[i] <- t; r <- .gen2[.gen2$Genus == t, ]; if (nrow(r)) { nd$Subfamily[i] <- r$Subfamily[1]; nd$Family[i] <- r$Family[1] } }
    if (l == "Species_bin") { nd$Species_bin[i] <- t; r <- .sp2[.sp2$Species_bin == t, ]; if (nrow(r)) { nd$Genus[i] <- r$Genus[1]; nd$Subfamily[i] <- r$Subfamily[1]; nd$Family[i] <- r$Family[1] } } }
  nd$key <- paste(nd$lev, nd$taxon, sep = ":"); nd$stars <- .pstar(nd$p)
  nd$dir <- ifelse(nd$est > 0, "Away from lot", "Toward lot")
  haskey <- function(lv2, t) if (is.na(t)) NA else { k <- paste0(lv2, ":", t); if (k %in% nd$key) k else NA }
  nd$parent <- NA_character_
  for (i in seq_len(nrow(nd))) { l <- nd$lev[i]
    anc <- switch(l,
      Species_bin = c(haskey("Genus", nd$Genus[i]), haskey("Subfamily", nd$Subfamily[i]), haskey("Family", nd$Family[i])),
      Genus = c(haskey("Subfamily", nd$Subfamily[i]), haskey("Family", nd$Family[i])),
      Subfamily = c(haskey("Family", nd$Family[i])), Family = NA_character_)
    anc <- anc[!is.na(anc)]; if (length(anc)) nd$parent[i] <- anc[1] }
  list(nd = nd, curves = curves)
}

.SPX <- 1.6; .BOXW <- 1.15; .BOXH <- 0.56
.PAL_D <- c("Away from lot" = "#2C7FB8", "Toward lot" = "#D95F02")
hier_layout <- function(nd) {
  fam <- unique(na.omit(nd$Family)); sub <- unique(na.omit(nd$Subfamily))
  gen <- unique(na.omit(nd$Genus)); spp <- unique(na.omit(nd$Species_bin))
  nd$k1 <- match(nd$Family, fam); nd$k2 <- match(nd$Subfamily, sub); nd$k3 <- match(nd$Genus, gen); nd$k4 <- match(nd$Species_bin, spp)
  nd <- nd %>% mutate(across(c(k1, k2, k3, k4), ~tidyr::replace_na(., 0)))
  nd <- nd[order(nd$k1, nd$k2, nd$k3, nd$k4), ]
  ch <- split(nd$key, nd$parent); env <- new.env(); env$cnt <- 0; env$x <- setNames(rep(NA_real_, nrow(nd)), nd$key)
  assign_x <- function(k) { kids <- ch[[k]]
    if (is.null(kids)) { env$cnt <- env$cnt + 1; env$x[k] <- env$cnt }
    else { kids <- kids[order(match(kids, nd$key))]; for (c in kids) assign_x(c); env$x[k] <- mean(env$x[kids]) } }
  roots <- nd$key[is.na(nd$parent)]; roots <- roots[order(match(roots, nd$key))]; for (r in roots) assign_x(r)
  nd$x <- env$x[nd$key] * .SPX; nd
}
hier_plot <- function(res, title) {
  nd <- res$nd; curves <- res$curves; tiers <- data.frame(y = 4:1)
  if (nrow(nd) == 0) return(ggplot() + annotate("text", 0, 0, label = "No significant taxa") + theme_void() + ggtitle(title))
  nd <- hier_layout(nd)
  cv <- curves %>% semi_join(nd %>% select(lev, taxon), by = c("lev", "taxon")) %>%
    left_join(nd %>% select(lev, taxon, key, nx = x, ny = y, dir), by = c("lev", "taxon")) %>%
    group_by(key) %>% mutate(lo = min(lower, na.rm = TRUE), hi = max(upper, na.rm = TRUE),
      rng = ifelse(hi - lo <= 0, 1, hi - lo), xr = (dist_km - min(dist_km)) / (max(dist_km) - min(dist_km)),
      X = nx + (xr - 0.5) * .BOXW, Yl = ny + ((response - lo) / rng - 0.5) * .BOXH,
      Ylo = ny + ((lower - lo) / rng - 0.5) * .BOXH, Yhi = ny + ((upper - lo) / rng - 0.5) * .BOXH) %>%
    ungroup() %>% arrange(key, X)
  box <- nd %>% transmute(key, dir, xmin = x - .BOXW/2, xmax = x + .BOXW/2, ymin = y - .BOXH/2, ymax = y + .BOXH/2)
  seg <- nd %>% filter(!is.na(parent)) %>% left_join(nd %>% select(parent = key, px = x, py = y), by = "parent")
  lab <- nd %>% mutate(italic = ifelse(lev %in% c("Genus", "Species_bin"), "italic", "plain"),
                       lab2 = paste0(taxon, " ", stars), ly = y - .BOXH/2 - 0.11)
  ggplot() +
    geom_hline(data = tiers, aes(yintercept = y), color = "#F2F2F2", linewidth = 10) +
    { if (nrow(seg)) geom_segment(data = seg, aes(x = px, y = py, xend = x, yend = y), color = "#C4C4C4", linewidth = 0.45) } +
    geom_rect(data = box, aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax, color = dir), fill = "white", linewidth = 0.5) +
    geom_ribbon(data = cv, aes(x = X, ymin = Ylo, ymax = Yhi, group = key, fill = dir), alpha = 0.22) +
    geom_line(data = cv, aes(x = X, y = Yl, group = key, color = dir), linewidth = 0.7) +
    geom_text(data = lab, aes(x = x, y = ly, label = lab2, fontface = italic), size = 2.7, vjust = 1, color = "#222222") +
    scale_color_manual(values = .PAL_D, name = "More active:", drop = FALSE) +
    scale_fill_manual(values = .PAL_D, name = "More active:", drop = FALSE) +
    scale_y_continuous(breaks = 4:1, labels = c("Family", "Subfamily", "Genus", "Species"), limits = c(0.3, 4.7)) +
    coord_cartesian(clip = "off") +
    labs(title = title, subtitle = "each inset = model-predicted detections vs distance (0–max km, auto-scaled) ± 95% CI; nodes link to nearest significant ancestor") +
    theme_void(base_family = "Arial") +
    theme(plot.title = element_text(face = "bold", size = 14), plot.subtitle = element_text(size = 9, color = "#555555"),
          axis.text.y = element_text(size = 10, face = "bold", color = "#333333", hjust = 1), legend.position = "top",
          legend.title = element_text(face = "bold", size = 9), legend.text = element_text(size = 9), plot.margin = margin(8, 14, 8, 8))
}

for (sc in list(list(drop = NULL, sfx = "", ttl = "Full data (all 14 sites)"),
                list(drop = far3, sfx = "_nofar", ttl = "Far sites removed (11 near sites)"))) {
  res <- hier_nodes(sc$drop)
  p <- hier_plot(res, paste0("Moth significant-taxa hierarchy (distance) — ", sc$ttl))
  nleaf <- if (nrow(res$nd)) sum(!(res$nd$key %in% res$nd$parent)) else 1
  ggsave(sprintf("output/figures/moth_sig_hierarchy_distance%s.png", sc$sfx), p,
         width = max(13, nleaf * 1.9), height = 8.5, dpi = 300, bg = "white", limitsize = FALSE)
  print(p)
}
```

### Bayesian distance hierarchy (brms, partial pooling)

A Bayesian counterpart to the hierarchy above. A single negative-binomial model **partially pools the distance slope up the taxonomy** — `(1 + dist_c | Family/Subfamily/Genus/Species)` — so each taxon's slope borrows strength from its relatives and sparsely-sampled taxa are regularised toward their group mean (no per-taxon multiple testing). "Significance" becomes the posterior **probability of direction**, P(dir) ≥ 0.95 (`*` ≥0.95, `**` ≥0.975, `***` ≥0.99). The model is fit on all species with ≥5 site-nights but, as above, only taxa with ≥30 site-nights of their own data are displayed.

Fits are cached under `output/brms_fits/`: the first knit fits three models (~15 min each via cmdstanr), later knits reuse the cache. Delete the cache files to force a refit. Requires `brms` + a working `cmdstan`; if unavailable the section is skipped.

```{r moth_brms_hierarchy, warning=FALSE, message=FALSE, fig.width=12, fig.height=7.5}
.brms_ok <- requireNamespace("brms", quietly = TRUE) &&
  requireNamespace("cmdstanr", quietly = TRUE) &&
  !is.null(tryCatch(cmdstanr::cmdstan_version(), error = function(e) NULL))

if (!.brms_ok) {
  message("brms/cmdstan not available — skipping Bayesian hierarchy. Install brms + cmdstan to enable.")
} else {
  suppressMessages({ library(brms); library(posterior) })
  dir.create("output/brms_fits", showWarnings = FALSE, recursive = TRUE)

  # species-level zero-filled grid (>=5 site-nights) with full lineage + scaled covariates
  brms_build_grid <- function(drop_sites = NULL) {
    ev <- insect_events %>%
      transmute(site, date = as.character(as.Date(date)), color, intensity,
                dist_km, jd, mean_phase, pct_nonforest, brightness_dark)
    spp <- sp_tax %>%
      filter(!is.na(Species_bin), !is.na(Family), !is.na(Subfamily), !is.na(Genus)) %>%
      mutate(date = as.character(as.Date(date)))
    if (!is.null(drop_sites)) { ev <- ev %>% filter(!site %in% drop_sites)
                                spp <- spp %>% filter(!site %in% drop_sites) }
    keep <- spp %>% group_by(Species_bin) %>%
      summarise(sn = n_distinct(paste(site, date)), .groups = "drop") %>%
      filter(sn >= 5) %>% pull(Species_bin)
    lin <- spp %>% filter(Species_bin %in% keep) %>%
      distinct(Species_bin, Genus, Subfamily, Family) %>%
      group_by(Species_bin) %>% slice(1) %>% ungroup()
    det <- spp %>% filter(Species_bin %in% keep) %>%
      group_by(site, date, color, intensity, Species_bin) %>%
      summarise(detections = n(), .groups = "drop")
    grid <- ev %>% tidyr::crossing(Species_bin = keep) %>%
      left_join(det, by = c("site","date","color","intensity","Species_bin")) %>%
      mutate(detections = tidyr::replace_na(detections, 0)) %>%
      left_join(lin, by = "Species_bin") %>%
      mutate(across(c(Family, Subfamily, Genus, Species_bin, site, color), factor),
             intensity = factor(intensity, levels = c("10","30","50","70","100")),
             jd_c = jd - mean(jd, na.rm = TRUE),
             dist_c = dist_km - mean(dist_km, na.rm = TRUE),
             mphase_z = as.numeric(scale(mean_phase)),
             nonf_z   = as.numeric(scale(pct_nonforest)),
             bdark_z  = as.numeric(scale(brightness_dark)))
    grid
  }

  brms_fit_hier <- function(grid, cache) {
    form <- bf(detections ~ dist_c + color + intensity + mphase_z + nonf_z + bdark_z +
                 jd_c + I(jd_c^2) + (1 | site) +
                 (1 + dist_c | Family/Subfamily/Genus/Species_bin))
    pri <- c(prior(normal(0, 1), class = b), prior(exponential(1), class = sd),
             prior(lkj(2), class = cor))
    brm(form, data = grid, family = negbinomial(), prior = pri,
        chains = 4, iter = 2000, warmup = 1000, cores = 4, backend = "cmdstanr",
        adapt_delta = 0.9, max_treedepth = 10, seed = 1, refresh = 0, file = cache)
  }

  # extract partially-pooled slopes + posterior-predicted curves; draw the tree
  brms_hier_plot <- function(fit, grid, title, min_sn = 30, pthr = 0.95) {
    lin <- grid %>% distinct(Family, Subfamily, Genus, Species_bin) %>%
      mutate(across(everything(), as.character))
    dr <- posterior::as_draws_df(fit)
    getc <- function(nm) if (nm %in% names(dr)) dr[[nm]] else rep(0, nrow(dr))
    b0 <- dr[["b_dist_c"]]; a0 <- dr[["b_Intercept"]]
    dgc <- seq(min(grid$dist_c), max(grid$dist_c), length.out = 40)
    cumab <- function(lev, f, s, g, sp) {
      a <- a0 + getc(sprintf("r_Family[%s,Intercept]", f))
      b <- b0 + getc(sprintf("r_Family[%s,dist_c]", f))
      if (lev %in% c("Subfamily","Genus","Species_bin")) {
        a <- a + getc(sprintf("r_Family:Subfamily[%s_%s,Intercept]", f, s))
        b <- b + getc(sprintf("r_Family:Subfamily[%s_%s,dist_c]", f, s)) }
      if (lev %in% c("Genus","Species_bin")) {
        a <- a + getc(sprintf("r_Family:Subfamily:Genus[%s_%s_%s,Intercept]", f, s, g))
        b <- b + getc(sprintf("r_Family:Subfamily:Genus[%s_%s_%s,dist_c]", f, s, g)) }
      if (lev == "Species_bin") {
        a <- a + getc(sprintf("r_Family:Subfamily:Genus:Species_bin[%s_%s_%s_%s,Intercept]", f, s, g, sp))
        b <- b + getc(sprintf("r_Family:Subfamily:Genus:Species_bin[%s_%s_%s_%s,dist_c]", f, s, g, sp)) }
      list(a = a, b = b)
    }
    lvtab <- list(
      Family      = lin %>% distinct(Family) %>% mutate(S = NA, G = NA, SP = NA, taxon = Family),
      Subfamily   = lin %>% distinct(Family, Subfamily) %>% mutate(G = NA, SP = NA, taxon = Subfamily) %>% rename(S = Subfamily),
      Genus       = lin %>% distinct(Family, Subfamily, Genus) %>% mutate(SP = NA, taxon = Genus) %>% rename(S = Subfamily, G = Genus),
      Species_bin = lin %>% mutate(taxon = Species_bin) %>% rename(S = Subfamily, G = Genus, SP = Species_bin))
    yy <- c(Family = 4, Subfamily = 3, Genus = 2, Species_bin = 1)
    nodes <- list(); curves <- list()
    for (lev in names(lvtab)) { tb <- lvtab[[lev]]
      for (i in seq_len(nrow(tb))) {
        f <- tb$Family[i]; s <- tb$S[i]; g <- tb$G[i]; sp <- tb$SP[i]; tx <- tb$taxon[i]
        ab <- cumab(lev, f, s, g, sp); b <- ab$b; pdir <- mean(b > 0); pd <- max(pdir, 1 - pdir)
        nodes[[length(nodes)+1]] <- tibble(lev = lev, taxon = tx, y = yy[lev],
          Family = f, Subfamily = s, Genus = g, Species_bin = sp,
          slope = median(b), lo = quantile(b, .025), hi = quantile(b, .975), pdir = pdir, pd = pd)
        mu <- exp(outer(ab$b, dgc) + ab$a)
        curves[[length(curves)+1]] <- tibble(lev = lev, taxon = tx, xx = dgc,
          response = apply(mu, 2, median), lower = apply(mu, 2, quantile, .025), upper = apply(mu, 2, quantile, .975))
      } }
    ND <- bind_rows(nodes); CV <- bind_rows(curves)
    pres <- grid %>% filter(detections > 0)
    snlk <- bind_rows(
      pres %>% group_by(taxon = Family)      %>% summarise(sn = n_distinct(paste(site, date)), .groups="drop") %>% mutate(lev = "Family"),
      pres %>% group_by(taxon = Subfamily)   %>% summarise(sn = n_distinct(paste(site, date)), .groups="drop") %>% mutate(lev = "Subfamily"),
      pres %>% group_by(taxon = Genus)       %>% summarise(sn = n_distinct(paste(site, date)), .groups="drop") %>% mutate(lev = "Genus"),
      pres %>% group_by(taxon = Species_bin) %>% summarise(sn = n_distinct(paste(site, date)), .groups="drop") %>% mutate(lev = "Species_bin")) %>%
      mutate(taxon = as.character(taxon))
    ND <- ND %>% left_join(snlk, by = c("lev","taxon")) %>% filter(sn >= min_sn, pd >= pthr)
    if (nrow(ND) == 0)
      return(ggplot() + annotate("text", 0, 0, label = "No credibly-directional taxa") + theme_void() + ggtitle(title))
    ND$dir <- ifelse(ND$slope > 0, "Away from lot", "Toward lot")
    ND$stars <- ifelse(ND$pd >= .99, "***", ifelse(ND$pd >= .975, "**", "*"))
    ND$key <- paste(ND$lev, ND$taxon, sep = ":")
    hk <- function(l, t) if (is.na(t)) NA else { k <- paste0(l, ":", t); if (k %in% ND$key) k else NA }
    ND$parent <- NA_character_
    for (i in seq_len(nrow(ND))) { l <- ND$lev[i]
      anc <- switch(l,
        Species_bin = c(hk("Genus", ND$Genus[i]), hk("Subfamily", ND$Subfamily[i]), hk("Family", ND$Family[i])),
        Genus = c(hk("Subfamily", ND$Subfamily[i]), hk("Family", ND$Family[i])),
        Subfamily = c(hk("Family", ND$Family[i])), Family = NA)
      anc <- anc[!is.na(anc)]; if (length(anc)) ND$parent[i] <- anc[1] }
    SPX <- 1.6; BOXW <- 1.15; BOXH <- 0.56; PAL <- c("Away from lot" = "#2C7FB8", "Toward lot" = "#D95F02")
    fam <- unique(na.omit(ND$Family)); sub <- unique(na.omit(ND$Subfamily)); gen <- unique(na.omit(ND$Genus)); spl <- unique(na.omit(ND$Species_bin))
    ND$k1 <- match(ND$Family, fam); ND$k2 <- match(ND$Subfamily, sub); ND$k3 <- match(ND$Genus, gen); ND$k4 <- match(ND$Species_bin, spl)
    ND <- ND %>% mutate(across(c(k1, k2, k3, k4), ~tidyr::replace_na(., 0))); ND <- ND[order(ND$k1, ND$k2, ND$k3, ND$k4), ]
    ch <- split(ND$key, ND$parent); env <- new.env(); env$cnt <- 0; env$x <- setNames(rep(NA_real_, nrow(ND)), ND$key)
    ax <- function(k) { kids <- ch[[k]]
      if (is.null(kids)) { env$cnt <- env$cnt + 1; env$x[k] <- env$cnt }
      else { kids <- kids[order(match(kids, ND$key))]; for (c in kids) ax(c); env$x[k] <- mean(env$x[kids]) } }
    roots <- ND$key[is.na(ND$parent)]; roots <- roots[order(match(roots, ND$key))]; for (r in roots) ax(r)
    ND$x <- env$x[ND$key] * SPX
    cv <- CV %>% semi_join(ND %>% select(lev, taxon), by = c("lev","taxon")) %>%
      left_join(ND %>% select(lev, taxon, key, nx = x, ny = y, dir), by = c("lev","taxon")) %>%
      group_by(key) %>% mutate(lo2 = min(lower), hi2 = max(upper), rng = ifelse(hi2 - lo2 <= 0, 1, hi2 - lo2),
        xr = (xx - min(xx)) / (max(xx) - min(xx)), X = nx + (xr - 0.5) * BOXW,
        Yl = ny + ((response - lo2) / rng - 0.5) * BOXH,
        Ylo = ny + ((lower - lo2) / rng - 0.5) * BOXH, Yhi = ny + ((upper - lo2) / rng - 0.5) * BOXH) %>%
      ungroup() %>% arrange(key, X)
    box <- ND %>% transmute(key, dir, xmin = x - BOXW/2, xmax = x + BOXW/2, ymin = y - BOXH/2, ymax = y + BOXH/2)
    seg <- ND %>% filter(!is.na(parent)) %>% left_join(ND %>% select(parent = key, px = x, py = y), by = "parent")
    lab <- ND %>% mutate(italic = ifelse(lev %in% c("Genus","Species_bin"), "italic", "plain"),
                         lab2 = paste0(taxon, " ", stars), ly = y - BOXH/2 - 0.12)
    tiers <- data.frame(y = 4:1)
    ggplot() +
      geom_hline(data = tiers, aes(yintercept = y), color = "#F2F2F2", linewidth = 10) +
      { if (nrow(seg)) geom_segment(data = seg, aes(x = px, y = py, xend = x, yend = y), color = "#C4C4C4", linewidth = 0.45) } +
      geom_rect(data = box, aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax, color = dir), fill = "white", linewidth = 0.5) +
      geom_ribbon(data = cv, aes(x = X, ymin = Ylo, ymax = Yhi, group = key, fill = dir), alpha = 0.22) +
      geom_line(data = cv, aes(x = X, y = Yl, group = key, color = dir), linewidth = 0.7) +
      geom_text(data = lab, aes(x = x, y = ly, label = lab2, fontface = italic), size = 2.7, vjust = 1, color = "#222222") +
      scale_color_manual(values = PAL, name = "More active:", drop = FALSE) +
      scale_fill_manual(values = PAL, name = "More active:", drop = FALSE) +
      scale_y_continuous(breaks = 4:1, labels = c("Family","Subfamily","Genus","Species"), limits = c(0.3, 4.7)) +
      coord_cartesian(clip = "off") +
      labs(title = title,
           subtitle = "Nodes = taxa with P(direction) ≥ 0.95 (* ≥0.95, ** ≥0.975, *** ≥0.99); inset = posterior-predicted detections vs distance (auto-scaled) ±95% CrI") +
      theme_void(base_family = "Arial") +
      theme(plot.title = element_text(face = "bold", size = 14), plot.subtitle = element_text(size = 8.5, color = "#555555"),
            axis.text.y = element_text(size = 10, face = "bold", color = "#333333", hjust = 1), legend.position = "top",
            legend.title = element_text(face = "bold", size = 9), legend.text = element_text(size = 9), plot.margin = margin(8, 14, 8, 8))
  }

  brms_scn <- list(
    list(drop = NULL,                              sfx = "full",   ttl = "full data, all 14 sites"),
    list(drop = c("GRTE13","GRTE14","GRTE15"),     sfx = "nofar",  ttl = "far sites removed, 11 near sites"),
    list(drop = "GRTE15",                          sfx = "site15", ttl = "GRTE15 removed"))
  for (sc in brms_scn) {
    g  <- brms_build_grid(sc$drop)
    ft <- brms_fit_hier(g, sprintf("output/brms_fits/moth_hier_%s", sc$sfx))
    pp <- brms_hier_plot(ft, g, paste0("Moth distance hierarchy — brms partially-pooled slopes (", sc$ttl, ")"))
    ggsave(sprintf("output/figures/moth_sig_hierarchy_brms_%s.png", sc$sfx), pp,
           width = 12, height = 7.5, dpi = 300, bg = "white", limitsize = FALSE)
    print(pp)
  }
}
```

### Moth color × intensity, by subfamily

Subfamily-resolution counterpart to the overall color × intensity model (storyboard panel 2.7). Each facet is one well-sampled subfamily; the White-vs-Red contrast is held common across subfamilies and marked at each intensity.

```{r insect_subfamily_colorintensity, warning=FALSE, message=FALSE, fig.width=10, fig.height=6.5}
# Color x intensity marginal means at subfamily resolution (full + far-removed).
make_taxon_ci <- function(level, out_tag, pretty, exclude_rx = NULL,
                          n_top = 9, min_sitenights = 30, drop_sites = NULL) {
  sfx <- if (is.null(drop_sites)) "" else "_nofar"
  det <- sp_tax %>% filter(!is.na(.data[[level]])) %>% rename(taxon = all_of(level))
  if (!is.null(exclude_rx)) det <- det %>% filter(!str_detect(taxon, exclude_rx))
  keep <- det %>% group_by(taxon) %>%
    summarise(d = n(), sn = n_distinct(paste(site, date)), .groups = "drop") %>%
    filter(sn >= min_sitenights) %>% slice_max(d, n = n_top) %>% arrange(desc(d)) %>%
    pull(taxon)

  det_kept <- det %>% filter(taxon %in% keep) %>%
    group_by(site, date, color, intensity, taxon) %>%
    summarise(detections = n(), .groups = "drop")

  grid <- event_covariates %>%
    mutate(jd_c = jd - mean(jd, na.rm = TRUE)) %>%
    tidyr::crossing(taxon = keep) %>%
    left_join(det_kept, by = c("site","date","color","intensity","taxon")) %>%
    mutate(detections = tidyr::replace_na(detections, 0),
           taxon = factor(taxon, levels = keep), site = factor(site),
           color = factor(color),
           intensity = factor(intensity, levels = c("10","30","50","70","100")))
  if (!is.null(drop_sites))
    grid <- grid %>% filter(!site %in% drop_sites) %>% droplevels()

  stable <- function(m) !is.null(m) && isTRUE(m$sdr$pdHess) &&
    !any(is.na(suppressWarnings(sqrt(diag(vcov(m)$cond)))))
  # Progressively simpler mean structures until SEs are estimable. Fine
  # taxonomic resolution on the far-removed (11-site) data can over-parameterise;
  # habitat covariates and the date quadratic are the first to drop.
  cand <- c(
    "mean_phase + pct_nonforest + brightness_dark + dist_km + jd_c + I(jd_c^2)",
    "mean_phase + brightness_dark + dist_km + jd_c + I(jd_c^2)",
    "mean_phase + dist_km + jd_c + I(jd_c^2)",
    "mean_phase + dist_km + jd_c",
    "dist_km + jd_c")
  m <- NULL
  for (rhs in cand) {
    mi <- tryCatch(suppressWarnings(glmmTMB(
      as.formula(paste("detections ~ taxon * intensity + color * intensity +",
                       rhs, "+ (1 | site)")),
      data = grid, ziformula = ~1, family = nbinom2())), error = function(e) NULL)
    if (is.null(m) && !is.null(mi)) m <- mi   # keep last fitted as a floor
    if (stable(mi)) { m <- mi; break }
  }

  # White-vs-Red ratio at each intensity (common across taxa)
  cc <- as.data.frame(contrast(emmeans(m, ~ color | intensity, type = "response"),
                               method = "revpairwise")) %>%
    mutate(stars = ifelse(p.value < .001, "***",
                   ifelse(p.value < .01, "**",
                   ifelse(p.value < .05, "*", ""))),
           intensity = factor(intensity, levels = c("10","30","50","70","100")))
  cat("\n===", pretty, "White/Red ratio by intensity ===\n")
  print(cc %>% transmute(intensity, ratio = round(ratio, 2),
                         p = signif(p.value, 3), stars))

  e <- as.data.frame(emmeans(m, ~ color * intensity | taxon, type = "response"))
  names(e)[names(e) %in% c("response","rate","emmean")] <- "est"
  e <- e %>% mutate(taxon = factor(taxon, levels = keep))

  fig <- ggplot(e, aes(factor(intensity), est, color = color, group = color)) +
    geom_line(linewidth = 0.6, position = position_dodge(0.3)) +
    geom_point(size = 1.6, position = position_dodge(0.3)) +
    geom_errorbar(aes(ymin = asymp.LCL, ymax = asymp.UCL), width = 0.15,
                  position = position_dodge(0.3)) +
    geom_text(data = cc, aes(x = intensity, y = Inf, label = stars),
              inherit.aes = FALSE, vjust = 1.3, size = 3.4, color = "#111111") +
    scale_color_manual(values = c(R = col_red, W = col_white),
                       labels = c(R = "Red", W = "White"), name = "Light color") +
    scale_y_continuous(expand = expansion(mult = c(0.05, 0.15))) +
    facet_wrap(~ taxon, scales = "free_y") +
    labs(title = paste0("Moth Detections by Light Color × Intensity, by ", pretty),
         subtitle = if (is.null(drop_sites))
           "Marginal means ± 95% CI; asterisks: White vs Red (common across taxa)" else
           "Far sites (GRTE13/14/15) removed; asterisks: White vs Red (common across taxa)",
         x = "Intensity (%)", y = "Predicted detections per site-night") +
    bat_theme + theme(strip.text = element_text(face = "bold.italic", size = 11),
                      legend.position = "top")
  ggsave(sprintf("output/figures/moth_%s_colorintensity_panels%s.png", out_tag, sfx),
         fig, width = 10, height = 6.5, dpi = 300, bg = "white")
  print(fig)
}

make_taxon_ci("Subfamily", "subfamily", "Subfamily")
make_taxon_ci("Subfamily", "subfamily", "Subfamily", drop_sites = far3)
```

```{r insect_community, warning=FALSE, message=FALSE}
source("R/10_community_analysis.R")
moth_community <- run_community_analysis(
  insect_family_env, bat_theme, col_red, col_white,
  taxon = "Moth", unit = "Family"
)

moth_community$perm
moth_community$disp_test
moth_community$db_global
moth_community$db_margin
moth_community$mv_anova
```

```{r insect_community_figs, warning=FALSE, message=FALSE}
moth_community$fig_c1
moth_community$fig_c2
moth_community$fig_c3

ggsave("output/figures/moth_fig_dbrda.png", moth_community$fig_c1,
       width = 7, height = 6, dpi = 300, bg = "white")
ggsave("output/figures/moth_fig_community_heatmap.png", moth_community$fig_c2,
       width = 8, height = 6, dpi = 300, bg = "white")
ggsave("output/figures/moth_fig_dist_heatmap.png", moth_community$fig_c3,
       width = 7, height = 6, dpi = 300, bg = "white")
```

## Moth Taxonomic Levels & Richness

Extends the family-level moth analysis to **subfamily, genus, and species** resolution. For each level we test whether light color, intensity, and distance restructure the community (dbRDA / PERMANOVA / mvabund), and we model **taxonomic richness** per site-night (species / genus / subfamily), adjusted for catch size. `mvabund` uses fewer bootstraps at the higher-dimensional levels (159 genera, 253 species) to keep knit time reasonable.

```{r moth_taxa_env, warning=FALSE, message=FALSE}
source("R/33_moth_taxonomy.R")
subfam_env  <- build_taxon_env(insects$specimens, event_covariates, "Subfamily")
genus_env   <- build_taxon_env(insects$specimens, event_covariates, "Genus")
species_env <- build_taxon_env(insects$specimens, event_covariates, "Species_bin")
c(subfamilies = n_distinct(subfam_env$species),
  genera      = n_distinct(genus_env$species),
  species     = n_distinct(species_env$species))
```

```{r moth_taxa_community, warning=FALSE, message=FALSE}
comm_subfam  <- run_community_analysis(subfam_env,  bat_theme, col_red, col_white,
                                       taxon = "Moth", unit = "Subfamily")
comm_genus   <- run_community_analysis(genus_env,   bat_theme, col_red, col_white,
                                       taxon = "Moth", unit = "Genus",   nBoot = 999)
comm_species <- run_community_analysis(species_env, bat_theme, col_red, col_white,
                                       taxon = "Moth", unit = "Species", nBoot = 999)

cat("=== Subfamily community (dbRDA marginal) ===\n"); comm_subfam$db_margin
cat("=== Genus community (dbRDA marginal) ===\n");     comm_genus$db_margin
cat("=== Species community (dbRDA marginal) ===\n");   comm_species$db_margin
```

```{r moth_taxa_community_figs, warning=FALSE, message=FALSE}
comm_subfam$fig_c3   # subfamily x distance heatmap
comm_genus$fig_c3    # genus x distance heatmap
comm_species$fig_c3  # species x distance heatmap
```

```{r moth_richness, warning=FALSE, message=FALSE}
moth_richness <- compute_richness(insects$specimens, event_covariates)
richness_results <- run_richness_models(moth_richness, bat_theme, intensity_pal,
                                        col_red, col_white)
richness_results$fig_richness
```

## Bat vs. Moth Comparison

```{r bat_moth_comparison, warning=FALSE, message=FALSE}
# Pull comparable marginal effects from each taxon's general model
bat_color_int <- as.data.frame(emmeans(simple_model1, ~ color | intensity, type = "response")) %>%
  rename(lower_CL = asymp.LCL, upper_CL = asymp.UCL) %>%
  mutate(taxon = "Bat")

moth_color_int <- as.data.frame(insect_general$em_color_intensity) %>%
  rename(lower_CL = asymp.LCL, upper_CL = asymp.UCL) %>%
  mutate(taxon = "Moth")

compare_df <- bind_rows(bat_color_int, moth_color_int) %>%
  mutate(intensity = factor(intensity, levels = c("10","30","50","70","100")))

fig_compare <- ggplot(compare_df,
                      aes(x = intensity, y = response,
                          color = color, group = interaction(color, taxon))) +
  geom_line(linewidth = 0.8, position = position_dodge(0.3)) +
  geom_point(size = 3, position = position_dodge(0.3)) +
  geom_errorbar(aes(ymin = lower_CL, ymax = upper_CL),
                width = 0.15, linewidth = 0.6, position = position_dodge(0.3)) +
  scale_color_manual(values = c(R = col_red, W = col_white),
                     labels = c(R = "Red", W = "White"), name = "Light color") +
  facet_wrap(~ taxon, scales = "free_y") +
  labs(
    title    = "Bat vs. Moth: Light Color × Intensity Effect",
    subtitle = "Marginal means ± 95% CI · each taxon's own negative binomial GLMM",
    x        = "Light intensity (%)",
    y        = "Predicted detections per night"
  ) +
  bat_theme +
  theme(legend.position = "top")

fig_compare

ggsave("output/figures/fig_bat_vs_moth_color_intensity.png",
       fig_compare, width = 10, height = 5, dpi = 300, bg = "white")

# Distance slope comparison
bat_dist_slope <- as.data.frame(emtrends(simple_model1, ~ intensity, var = "dist_km", type = "response")) %>%
  mutate(taxon = "Bat")
moth_dist_slope <- as.data.frame(insect_general$em_dist_trend) %>%
  mutate(taxon = "Moth")

cat("=== Distance slope (per km) by intensity — Bat vs. Moth ===\n")
bind_rows(bat_dist_slope, moth_dist_slope) %>%
  select(taxon, intensity, dist_km.trend, SE) %>%
  arrange(intensity, taxon)
```

## Bat vs. Moth Relative Effects

```{r bat_moth_effects, warning=FALSE, message=FALSE}
source("R/17_bat_moth_effects.R")
bm_effects <- compare_bat_moth_effects(
  simple_model1, insect_general$model, data_env, insect_total_env,
  bat_theme, col_navy, col_purple
)

bm_effects$fig_color
bm_effects$fig_intensity
bm_effects$fig_distance
bm_effects$fig_forest

bm_effects$forest_df
```

## Robustness / Sensitivity Analysis

The three far sites (\~7 km) sit in a large gap beyond the rest (0–2 km), and GRTE15 in particular is a high-detection outlier. Refit the key models — the overall bat model, the guild model, and the moth activity model — with these sites removed, and compare the colour and distance effects. Colour effects are expected to hold; the distance/guild slopes are the ones likely to hinge on the far sites.

```{r robustness_analysis, warning=FALSE, message=FALSE}
library(emmeans)

far_sites <- c("GRTE13", "GRTE14", "GRTE15")   # the three far (~7 km) sites
robust_scenarios <- list("Full data"      = character(0),
                         "Drop GRTE15"     = "GRTE15",
                         "Drop far sites"  = far_sites)

# White-vs-Red ratio (+ p) from any model with a `color` term
.color_wr <- function(m) {
  ct <- as.data.frame(contrast(emmeans(m, ~ color, type = "response"),
                               method = "revpairwise"))
  sprintf("%.2f (p=%.3f)", ct$ratio[1], ct$p.value[1])
}
# marginal distance slope (log scale) via emtrends; `by` gives per-guild slopes
.dist_slope <- function(m, by = NULL) {
  et <- as.data.frame(if (is.null(by)) emtrends(m, ~ 1, var = "dist_km")
                      else emtrends(m, reformulate(by), var = "dist_km"))
  if (is.null(by)) sprintf("%.2f", et$dist_km.trend[1])
  else setNames(round(et$dist_km.trend, 2), et[[by]])
}

robustness <- purrr::map_dfr(names(robust_scenarios), function(nm) {
  drop <- robust_scenarios[[nm]]
  keep <- function(df) droplevels(dplyr::filter(df, !site %in% drop))
  # refit each model on the reduced data (same formula/family via update())
  bat  <- update(simple_model1,        data = keep(site_night))
  gld  <- update(cluster_glmm,         data = keep(cluster_night))
  moth <- update(insect_general$model, data = keep(insect_total_env))
  gs   <- .dist_slope(gld, by = "cluster")
  tibble::tibble(
    Scenario            = nm,
    `Bat color (W/R)`   = .color_wr(bat),
    `Bat dist slope`    = .dist_slope(bat),
    `Guild color (W/R)` = .color_wr(gld),
    `Guild dist Low/Myev/High` = paste(gs["LowFreq"], gs["Myev"], gs["HighFreq"], sep = " / "),
    `Moth color (W/R)`  = .color_wr(moth),
    `Moth dist slope`   = .dist_slope(moth)
  )
})

knitr::kable(
  robustness,
  caption = paste("Sensitivity of colour and distance effects to removing the",
                  "high-leverage far sites. Colour W/R = White-vs-Red ratio;",
                  "distance slopes are on the log scale."))
```

## Lamp Visual Modeling: Moth vs. Bat Perception

```{r lamp_visual_model, warning=FALSE, message=FALSE}
source("R/24_lamp_visual_model.R")
lamp_vis <- run_lamp_visual_model(bat_theme = bat_theme, col_red = col_red, col_white = col_white)

lamp_vis$fig_moth
lamp_vis$fig_bat
lamp_vis$fig_moth_abs
lamp_vis$fig_bat_abs
```

------------------------------------------------------------------------

## Visitor Perception Survey (Factor–Cluster Analysis)

Exploratory factor analysis of 13 streetlight attitude items (1–5 Likert, 1 = "not at all true") from 166 visitors, followed by k-means clustering on the factor scores. `StreetlightCondition` (1 = red, 2 = white) and `Acceptable_Sky_Conditions` (1–7) are held out and used to profile the segments. Missing values (`999`) are median-imputed. The chunk prints the KMO/Bartlett suitability tests, factor loadings, cluster silhouettes, and cluster profiles.

```{r people_factor_cluster, warning=FALSE, message=FALSE}
source("R/25_people_factor_cluster.R")
```

```{r people_cluster_figure, warning=FALSE, message=FALSE}
p   # respondent clusters in factor space
```

## Visitor Perception Survey (PCA + Cronbach's α)

Principal components analysis of the same 13 items as a data-reduction step, following the standard motivation-segmentation recipe: assess suitability (Bartlett's test of sphericity, KMO \> 0.50), extract components with eigenvalue \> 1, assign items at a minimum loading of ≥ 0.40, eliminate cross-loaded items, check each component's reliability with Cronbach's α (≥ 0.65), and average the retained items into a single component score. The chunk prints suitability tests, rotated loadings, the item-assignment table, per-component α, and the component scores related to light condition and sky acceptability.

```{r people_pca, warning=FALSE, message=FALSE}
source("R/26_people_pca.R")
```

## Predicting Light Exposure from Attitudes (Logistic GLM)

Binomial GLM asking whether the reduced attitude scores discriminate red vs. white light exposure (outcome: `StreetlightCondition`, white = 1, red = 0). Predictors are the EFA factor scores and the PCA component scores from the two sections above, so this chunk must run after them. A plain GLM is used because the survey has no grouping variable for a random effect. **Direction caveat:** light condition was experimentally assigned and attitudes were measured after exposure, so this is a discriminant/association model — which responses distinguish the two groups — not evidence that attitudes cause exposure.

```{r people_exposure_glm, warning=FALSE, message=FALSE}
source("R/27_people_exposure_glm.R")
```
