---
title: "Distance Project"
author: "Nigel K Anderson"
format: html
editor: visual
---

```{r setup, include=FALSE}
library(tidyverse)

# Shared theme + palette (matches ColterBay style)
bat_theme <- theme_classic(base_size = 12) +
  theme(
    plot.background    = element_rect(fill = "white", color = NA),
    panel.background   = element_rect(fill = "white", color = NA),
    panel.grid.major.y = element_line(color = "#EBEBEB", linewidth = 0.4),
    panel.grid.major.x = element_blank(),
    axis.line          = element_line(color = "#444444", linewidth = 0.5),
    axis.ticks         = element_line(color = "#444444"),
    axis.text          = element_text(color = "#333333", size = 10),
    axis.title         = element_text(color = "#222222", size = 11),
    plot.title         = element_text(color = "#111111", size = 13,
                                      face = "bold", margin = margin(b = 3)),
    plot.subtitle      = element_text(color = "#666666", size = 9,
                                      margin = margin(b = 8)),
    legend.background  = element_rect(fill = "white", color = NA),
    legend.key         = element_rect(fill = "white", color = NA),
    legend.text        = element_text(color = "#333333", size = 10),
    legend.title       = element_text(color = "#444444", size = 9, face = "bold"),
    strip.background   = element_rect(fill = "#F2F2F2", color = NA),
    strip.text         = element_text(color = "#222222", face = "bold", size = 10),
    plot.caption       = element_text(color = "#999999", size = 8,
                                      hjust = 0, margin = margin(t = 8)),
    plot.margin        = margin(12, 14, 10, 12)
  )

col_red    <- "#7A1C2E"   
col_white  <- "#C9A84C"   
col_navy   <- "#1B2A4A"   
col_purple <- "#4A1259"   
col_ivory  <- "#F5F0E3"   

intensity_pal <- c("10"  = col_ivory,
                   "30"  = col_white,
                   "50"  = col_purple,
                   "70"  = col_red,
                   "100" = col_navy)
```

## Data Pipeline

```{r import, warning=FALSE, message=TRUE}
source("R/01_import_sonobat.R")
```

```{r clean, warning=FALSE, message=FALSE}
source("R/02_clean_data.R")
data <- clean_data(data_raw)
```

```{r habitat, warning=FALSE, message=FALSE}
source("R/03_extract_habitat.R")

data_env <- add_habitat(
  data,
  forest_raster = nlcd_forest,
  buffer = 50
)

data_env <- add_moonlight(data_env)
```

```{r treatment, warning=FALSE, message=FALSE}
source("R/07_add_light_treatment.R")

data_env <- data_env %>%
  left_join(
    lighting %>% select(date, color, intensity),
    by = "date"
  ) %>%
  mutate(
    color     = factor(color),
    site      = factor(site),
    intensity = factor(intensity)
  )

```

```{r distance, warning=FALSE, message=FALSE}
library(sf)

ref_point <- st_as_sf(
  data.frame(lon = -110.645134, lat = 43.904879),
  coords = c("lon", "lat"),
  crs = 4326
)

sites_sf <- st_as_sf(sites, coords = c("lon", "lat"), crs = 4326)
sites_sf <- st_transform(sites_sf, 32612)
ref_point <- st_transform(ref_point, 32612)

sites_sf$dist_m <- as.numeric(st_distance(sites_sf, ref_point))

# Pull only dist_m — avoid attaching the sf geometry class to data_env
site_dists <- as.data.frame(sites_sf) %>% dplyr::select(site, dist_m)

data_env <- data_env %>%
  left_join(site_dists, by = "site") %>%
  mutate(
    dist_km = dist_m / 1000,
    dist_cat = case_when(
      dist_km < 0.52 ~ "Close",
      dist_km < 1    ~ "Medium",
      dist_km < 2.1  ~ "Further",
      TRUE           ~ "Far"
    ),
    jd_c = jd - mean(jd, na.rm = TRUE)   # centered Julian day
  )

data_env <- data_env %>% filter(!is.na(dist_km))

data_env %>% select(site, dist_km) %>% distinct() %>% arrange(dist_km)

# Filter to focal species (matches ColterBay project)
focal_spp <- c("Laci", "Lano", "Mylu", "Epfu", "Myev", "Myvo")
data_env  <- data_env %>% filter(species %in% focal_spp)

message("Rows after species filter: ", nrow(data_env),
        "  |  Species retained: ", n_distinct(data_env$species))
```

```{r sky_brightness, warning=FALSE, message=TRUE}
source("R/09_add_sky_brightness.R")
```

```{r bat_traits, warning=FALSE, message=FALSE}
library(readxl)

bat_traits <- read_excel("data/bat_traits.xlsx") %>%
  rename(species = Species)

data_env <- data_env %>%
  left_join(bat_traits, by = "species")
```

## Models

```{r simple_model, warning=FALSE, message=FALSE}
library(glmmTMB)

simple_model1 <- glmmTMB(
  detections ~
    color * intensity +
    intensity * dist_km +
    mean_phase +
    pct_nonforest +
    brightness_dark +
    jd_c + I(jd_c^2) +
    (1 | site),
  data      = data_env,
  ziformula = ~1,
  family    = nbinom2()
)

summary(simple_model1)
```

## Community Analysis

```{r community, warning=FALSE, message=FALSE}
source("R/10_community_analysis.R")
community <- run_community_analysis(data_env, bat_theme, col_red, col_white)
```

```{r community_results, warning=FALSE, message=FALSE}
# NMDS stress
cat("NMDS stress:", round(community$nmds$stress, 3), "\n")

# PERMANOVA
community$perm

# Homogeneity of dispersion (by color)
community$disp_test

# dbRDA — global and marginal tests
community$db_global
community$db_margin

# mvabund — multivariate GLM
community$mv_anova
```

```{r community_figs, warning=FALSE, message=FALSE}
community$fig_c1   # dbRDA ordination biplot
community$fig_c2   # Species activity heatmap (color × intensity)
community$fig_c3   # Species activity heatmap (distance)
```

## Diagnostics

```{r diagnostics, warning=FALSE, message=FALSE}
library(corrplot)

# Covariate correlation matrix
vars <- data_env %>%
  select(detections, pct_nonforest, jd, mean_phase, dist_km)
corrplot(cor(vars, use = "complete.obs"))

# Site-level means
site_means <- data_env %>%
  group_by(site, dist_km) %>%
  summarise(mean_det = mean(detections, na.rm = TRUE), .groups = "drop")

# Scatter with smooth
ggplot(site_means, aes(x = dist_km, y = mean_det)) +
  geom_smooth(method = "lm", se = TRUE, color = "#2E86AB",
              fill = "#DDEAF5", linewidth = 1) +
  geom_point(size = 3.5, color = "#333333") +
  ggrepel::geom_text_repel(aes(label = site), size = 3, color = "#666666") +
  labs(
    title    = "Bat Detections vs Distance from Light Source",
    subtitle = "Site-level means with linear trend",
    x        = "Distance from light (km)",
    y        = "Mean detections per night"
  ) +
  bat_theme

# Correlation test
cor.test(site_means$dist_km, site_means$mean_det, method = "spearman")


# (Distance post-hoc contrasts moved to posthoc chunk below)
```

## Post-hoc Contrasts

```{r posthoc, warning=FALSE, message=FALSE}
library(emmeans)

# ── 1. Color × Intensity simple effects ───────────────────────────────────────
# Red vs. White at each intensity level (the key interaction question)
em_color_int <- emmeans(simple_model1, ~ color | intensity, type = "response")

cat("=== Estimated marginal means: color at each intensity ===\n")
em_color_int

cat("\n=== Red vs. White contrast at each intensity (Holm-corrected) ===\n")
pairs(em_color_int, adjust = "holm")

# ── 2. Intensity simple effects within each color ─────────────────────────────
# Which intensity levels differ from each other, separately for red and white?
em_int_color <- emmeans(simple_model1, ~ intensity | color, type = "response")

cat("\n=== Intensity pairwise contrasts within Red (Holm-corrected) ===\n")
pairs(em_int_color, adjust = "holm")

# ── 3. Intensity × Distance interaction test ──────────────────────────────────
em_dist_trend <- emtrends(simple_model1, ~ intensity,
                           var = "dist_km", type = "response")

cat("\n=== Distance slope (km) by intensity level ===\n")
em_dist_trend

cat("\n=== Pairwise intensity distance-slope contrasts (Holm-corrected) ===\n")
pairs(em_dist_trend, adjust = "holm")

# Contrasts at meaningful distances
em_dist_points <- emmeans(
  simple_model1,
  ~ intensity | dist_km,
  at   = list(dist_km = c(0.3, 0.7, 2.0)),
  type = "response"
)

cat("\n=== Intensity contrasts at close (0.3 km), medium (0.7 km), far (2.0 km) ===\n")
pairs(em_dist_points, adjust = "holm")
```

## Distance × Intensity Figure

```{r dist_intensity_fig, warning=FALSE, message=FALSE}
library(emmeans)

# Predicted detections across the observed distance gradient, by intensity
dist_range <- range(data_env$dist_km, na.rm = TRUE)
dist_grid  <- seq(dist_range[1], dist_range[2], length.out = 60)

em_dist_int <- emmeans(
  simple_model1,
  ~ intensity | dist_km,
  at      = list(dist_km = dist_grid),
  type    = "response"
)

pred_df <- as.data.frame(em_dist_int) %>%
  rename(lower_CL = asymp.LCL, upper_CL = asymp.UCL)

ggplot(pred_df, aes(x = dist_km, y = response,
                    color = factor(intensity), fill = factor(intensity))) +
  geom_ribbon(aes(ymin = lower_CL, ymax = upper_CL), alpha = 0.12, color = NA) +
  geom_line(linewidth = 1.1) +
  scale_color_manual(values = intensity_pal, name = "Intensity (%)") +
  scale_fill_manual(values  = intensity_pal, name = "Intensity (%)") +
  scale_x_continuous(breaks = scales::pretty_breaks(n = 5)) +
  labs(
    title    = "Predicted Bat Detections by Distance and Light Intensity",
    subtitle = "Marginal means ± 95% CI from negative binomial GLMM",
    x        = "Distance from Colter Bay parking lot (km)",
    y        = "Predicted detections per night",
    caption  = "Other covariates held at their means"
  ) +
  bat_theme +
  theme(legend.position = "right")

ggsave("output/figures/fig9_dist_intensity.png",
       last_plot(), width = 8, height = 5, dpi = 300, bg = "white")
```

## Morphology Trait Analysis

```{r trait_models, warning=FALSE, message=FALSE}
source("R/12_trait_models.R")
trait_results <- run_trait_models(data_env, bat_theme, col_navy, col_red, col_purple)

for (trait in names(trait_results$trait_models)) {
  cat("\n\n=== ", trait, " model ===\n")
  print(summary(trait_results$trait_models[[trait]]))
}
```

```{r trait_posthoc, warning=FALSE, message=FALSE}
cat("=== Intensity effect at low/mean/high trait values ===\n")
trait_results$df_trait_intensity

cat("\n=== Distance slope at low/mean/high trait values ===\n")
trait_results$df_trait_slope
```

```{r trait_figs, warning=FALSE, message=FALSE}
trait_results$fig_trait_intensity
trait_results$fig_trait_dist
```

## Species Models

```{r species_models, warning=FALSE, message=FALSE}
source("R/11_species_models.R")
sp_results <- run_species_models(data_env, bat_theme, col_red, col_white)
```

```{r species_figs, warning=FALSE, message=FALSE}
sp_results$fig_s1   # Color effect by species
sp_results$fig_s2   # Intensity slopes by species
sp_results$fig_s3   # White:red ratio heatmap
sp_results$fig_s4a  # Predicted detections ~ dist_km × color × intensity (species facets)
```

## Figures

```{r figures, message=FALSE, warning=FALSE}
source("R/08_figures.R")

fig1; fig2; fig3; fig4; fig5; fig6; fig7; fig8; fig9; fig11; fig12; fig13; fig14; fig15; fig16; fig19; fig20; fig21; fig22; fig23; fig24; fig25; fig26; fig27
```

## Species Key

| Code | Species                                          |
|------|--------------------------------------------------|
| Myyu | Yuma myotis (possibly Myvo)                      |
| Myvo | *Myotis volans* \*                               |
| Myth | *Myotis thysanodes* (possibly Myev)              |
| Mylu | *Myotis lucifugus* \*                            |
| Myev | *Myotis evotis* \*                               |
| Myci | *Myotis ciliolabrum*                             |
| Myca | *Myotis californicus*                            |
| Lano | *Lasionycteris noctivagans*                      |
| Laci | *Lasiurus cinereus* \*                           |
| Euma | *Euderma maculatum*                              |
| Epfu | *Eptesicus fuscus* \*                            |
| Anpa | *Antrozous pallidus* (possibly *L. noctivagans*) |

## Moth (Insect) Data Import

```{r insect_import, warning=FALSE, message=FALSE}
source("R/13_import_insects.R")
insects <- import_insects("data/grte_distance_insectID.xlsx")
```

```{r insect_env, warning=FALSE, message=FALSE}
insect_events <- insects$total %>%
  left_join(sites, by = "site") %>%
  left_join(site_dists, by = "site") %>%
  mutate(
    dist_km  = dist_m / 1000,
    dist_cat = case_when(
      dist_km < 0.52 ~ "Close",
      dist_km < 1    ~ "Medium",
      dist_km < 2.1  ~ "Further",
      TRUE           ~ "Far"
    ),
    jd = lubridate::yday(date)
  ) %>%
  filter(!is.na(dist_km))

insect_events <- add_habitat(insect_events, forest_raster = nlcd_forest, buffer = 50,
                              cache_path = "data/insect_data_out.rds")

# Photos are dated but not exactly timestamped; noon avoids any timezone date-rollover
insect_events$datetime <- as.POSIXct(paste(insect_events$date, "12:00:00"), tz = "America/Denver")
insect_events <- add_moonlight(insect_events)

sky <- read_csv("data/sky_brightness_measurements.csv", show_col_types = FALSE) %>%
  mutate(site = toupper(site))
insect_events <- insect_events %>%
  left_join(sky %>% select(site, brightness_dark), by = "site")

insect_total_env <- insect_events %>%
  mutate(
    jd_c      = jd - mean(jd, na.rm = TRUE),
    color     = factor(color),
    site      = factor(site),
    intensity = factor(intensity)
  )

message("Moth sampling events: ", nrow(insect_total_env),
        "  |  Sites: ", n_distinct(insect_total_env$site),
        "  |  Date range: ", paste(range(insect_total_env$date), collapse = " to "))
```

## Moth General Activity Model

```{r insect_general_model, warning=FALSE, message=FALSE}
source("R/15_insect_models.R")
insect_general <- run_insect_general_model(insect_total_env, bat_theme, intensity_pal)

summary(insect_general$model)
```

```{r insect_general_posthoc, warning=FALSE, message=FALSE}
cat("=== Color effect within each intensity level ===\n")
insect_general$em_color_intensity
pairs(insect_general$em_color_intensity, adjust = "holm")

cat("\n=== Distance slope by intensity level ===\n")
insect_general$em_dist_trend
```

```{r insect_general_figs, warning=FALSE, message=FALSE}
insect_general$fig_color_intensity
insect_general$fig_dist_intensity
```

```{r insect_moon_phase, warning=FALSE, message=FALSE}
source("R/22_moth_moon_phase.R")
fig_moth_moon_phase <- plot_moth_moon_phase(insect_total_env, bat_theme, col_red, col_white)
fig_moth_moon_phase
```

```{r insect_100red_diagnostic, warning=FALSE, message=FALSE}
source("R/23_moth_intensity_family.R")
diag_100red <- plot_moth_100red_diagnostic(insect_total_env, insects$specimens, bat_theme, col_red, col_white)
diag_100red$fig_combined
```

## Moth Community Analysis (Family)

```{r insect_family_env, warning=FALSE, message=FALSE}
event_covariates <- insect_events %>%
  select(site, date, color, intensity, dist_km, dist_cat, jd,
         mean_phase, mean_moonlight, max_moonlight,
         pct_forest, pct_nonforest, brightness_dark)

insect_family_env <- insects$family %>%
  left_join(event_covariates, by = c("site", "date", "color", "intensity")) %>%
  filter(!is.na(dist_km)) %>%
  rename(species = Family) %>%
  mutate(color = factor(color), site = factor(site), intensity = factor(intensity))

message("Moth family-level rows: ", nrow(insect_family_env),
        "  |  Families: ", paste(sort(unique(insect_family_env$species)), collapse = ", "))
```

```{r insect_family_models, warning=FALSE, message=FALSE}
source("R/11_species_models.R")
moth_family_results <- run_species_models(
  insect_family_env, bat_theme, col_red, col_white,
  taxon = "Moth", unit = "Family", file_prefix = "moth_"
)
```

```{r insect_family_figs, warning=FALSE, message=FALSE}
moth_family_results$fig_s1
moth_family_results$fig_s2
moth_family_results$fig_s3
moth_family_results$fig_s4a
```

```{r insect_family_site, warning=FALSE, message=FALSE}
source("R/21_moth_family_site.R")
fig_moth_family_site <- plot_moth_family_site(insect_family_env, bat_theme, col_red, col_navy, col_ivory)
fig_moth_family_site
```

```{r insect_community, warning=FALSE, message=FALSE}
source("R/10_community_analysis.R")
moth_community <- run_community_analysis(
  insect_family_env, bat_theme, col_red, col_white,
  taxon = "Moth", unit = "Family"
)

moth_community$perm
moth_community$disp_test
moth_community$db_global
moth_community$db_margin
moth_community$mv_anova
```

```{r insect_community_figs, warning=FALSE, message=FALSE}
moth_community$fig_c1
moth_community$fig_c2
moth_community$fig_c3

ggsave("output/figures/moth_fig_dbrda.png", moth_community$fig_c1,
       width = 7, height = 6, dpi = 300, bg = "white")
ggsave("output/figures/moth_fig_community_heatmap.png", moth_community$fig_c2,
       width = 8, height = 6, dpi = 300, bg = "white")
ggsave("output/figures/moth_fig_dist_heatmap.png", moth_community$fig_c3,
       width = 7, height = 6, dpi = 300, bg = "white")
```

## Bat vs. Moth Comparison

```{r bat_moth_comparison, warning=FALSE, message=FALSE}
# Pull comparable marginal effects from each taxon's general model
bat_color_int <- as.data.frame(emmeans(simple_model1, ~ color | intensity, type = "response")) %>%
  rename(lower_CL = asymp.LCL, upper_CL = asymp.UCL) %>%
  mutate(taxon = "Bat")

moth_color_int <- as.data.frame(insect_general$em_color_intensity) %>%
  rename(lower_CL = asymp.LCL, upper_CL = asymp.UCL) %>%
  mutate(taxon = "Moth")

compare_df <- bind_rows(bat_color_int, moth_color_int) %>%
  mutate(intensity = factor(intensity, levels = c("10","30","50","70","100")))

fig_compare <- ggplot(compare_df,
                      aes(x = intensity, y = response,
                          color = color, group = interaction(color, taxon))) +
  geom_line(linewidth = 0.8, position = position_dodge(0.3)) +
  geom_point(size = 3, position = position_dodge(0.3)) +
  geom_errorbar(aes(ymin = lower_CL, ymax = upper_CL),
                width = 0.15, linewidth = 0.6, position = position_dodge(0.3)) +
  scale_color_manual(values = c(R = col_red, W = col_white),
                     labels = c(R = "Red", W = "White"), name = "Light color") +
  facet_wrap(~ taxon, scales = "free_y") +
  labs(
    title    = "Bat vs. Moth: Light Color × Intensity Effect",
    subtitle = "Marginal means ± 95% CI · each taxon's own negative binomial GLMM",
    x        = "Light intensity (%)",
    y        = "Predicted detections per night"
  ) +
  bat_theme +
  theme(legend.position = "top")

fig_compare

ggsave("output/figures/fig_bat_vs_moth_color_intensity.png",
       fig_compare, width = 10, height = 5, dpi = 300, bg = "white")

# Distance slope comparison
bat_dist_slope <- as.data.frame(emtrends(simple_model1, ~ intensity, var = "dist_km", type = "response")) %>%
  mutate(taxon = "Bat")
moth_dist_slope <- as.data.frame(insect_general$em_dist_trend) %>%
  mutate(taxon = "Moth")

cat("=== Distance slope (per km) by intensity — Bat vs. Moth ===\n")
bind_rows(bat_dist_slope, moth_dist_slope) %>%
  select(taxon, intensity, dist_km.trend, SE) %>%
  arrange(intensity, taxon)
```

## Total Detections: Bat vs. Moth

```{r total_detections, warning=FALSE, message=FALSE}
source("R/16_total_detections.R")
totals <- plot_total_detections(
  data_env, insect_total_env, bat_theme,
  col_navy, col_purple, col_red, col_white, intensity_pal
)

totals$fig_site
totals$fig_color
totals$fig_intensity
```

## Bat vs. Moth Relative Effects

```{r bat_moth_effects, warning=FALSE, message=FALSE}
source("R/17_bat_moth_effects.R")
bm_effects <- compare_bat_moth_effects(
  simple_model1, insect_general$model, data_env, insect_total_env,
  bat_theme, col_navy, col_purple
)

bm_effects$fig_color
bm_effects$fig_intensity
bm_effects$fig_distance
bm_effects$fig_forest

bm_effects$forest_df
```

## Lamp Visual Modeling: Moth vs. Bat Perception

```{r lamp_visual_model, warning=FALSE, message=FALSE}
source("R/24_lamp_visual_model.R")
lamp_vis <- run_lamp_visual_model(bat_theme = bat_theme, col_red = col_red, col_white = col_white)

lamp_vis$fig_moth
lamp_vis$fig_bat
lamp_vis$fig_moth_abs
lamp_vis$fig_bat_abs
```

------------------------------------------------------------------------

## Occurrence Thresholds

```{r thresholds, warning=FALSE, message=FALSE}
source("R/04_Occurrence_Thresholds.R")

occurrences <- run_thresholds(data_env, thresholds = c(30, 60, 120, 300))

env_data <- data_env %>%
  distinct(site, date, dist_km, mean_phase,
           color, intensity, jd_c, pct_nonforest, brightness_dark)

occurrences <- occurrences %>%
  left_join(env_data, by = c("site", "date")) %>%
  left_join(data_env %>% distinct(site, pct_forest), by = "site")

data_30  <- occurrences %>% filter(threshold_sec == 30)
data_60  <- occurrences %>% filter(threshold_sec == 60)
data_120 <- occurrences %>% filter(threshold_sec == 120)
data_300 <- occurrences %>% filter(threshold_sec == 300)
```

```{r threshold_models, warning=FALSE, message=FALSE}
threshold_formula <- occurrences ~
  color * intensity +
  intensity * dist_km +
  pct_nonforest +
  jd_c +
  I(jd_c^2) +
  mean_phase +
  brightness_dark +
  (1 | site)

data_30_model  <- glmmTMB(threshold_formula, data = data_30, ziformula = ~1,  family = nbinom2())
data_60_model  <- glmmTMB(threshold_formula, data = data_60, ziformula = ~1,  family = nbinom2())
data_120_model <- glmmTMB(threshold_formula, data = data_120, ziformula = ~1, family = nbinom2())
data_300_model <- glmmTMB(threshold_formula, data = data_300, ziformula = ~1, family = nbinom2())

AIC(data_30_model, data_60_model, data_120_model, data_300_model)

summary(data_30_model)
```

```{r threshold_figures, warning=FALSE, message=FALSE}
source("R/08b_threshold_figures.R")

fig_t1; fig_t2; fig_t3; fig_t4; fig_t5; fig_t6; fig_t7; fig_t8; fig_t9; fig_t10; fig_t11; fig_t12; fig_t13; fig_t14; fig_t15; fig_t16; fig_t17
```

## Visitor Perception Survey (Factor–Cluster Analysis)

Exploratory factor analysis of 13 streetlight attitude items (1–5 Likert, 1 = "not at all true") from 166 visitors, followed by k-means clustering on the factor scores. `StreetlightCondition` (1 = red, 2 = white) and `Acceptable_Sky_Conditions` (1–7) are held out and used to profile the segments. Missing values (`999`) are median-imputed. The chunk prints the KMO/Bartlett suitability tests, factor loadings, cluster silhouettes, and cluster profiles.

```{r people_factor_cluster, warning=FALSE, message=FALSE}
source("R/25_people_factor_cluster.R")
```

```{r people_cluster_figure, warning=FALSE, message=FALSE}
p   # respondent clusters in factor space
```

## Visitor Perception Survey (PCA + Cronbach's α)

Principal components analysis of the same 13 items as a data-reduction step, following the standard motivation-segmentation recipe: assess suitability (Bartlett's test of sphericity, KMO \> 0.50), extract components with eigenvalue \> 1, assign items at a minimum loading of ≥ 0.40, eliminate cross-loaded items, check each component's reliability with Cronbach's α (≥ 0.65), and average the retained items into a single component score. The chunk prints suitability tests, rotated loadings, the item-assignment table, per-component α, and the component scores related to light condition and sky acceptability.

```{r people_pca, warning=FALSE, message=FALSE}
source("R/26_people_pca.R")
```
