---
title: "Distance Project"
author: "Nigel K Anderson"
format: html
editor: visual
---

```{r setup, include=FALSE}
library(tidyverse)

# Shared theme + palette (matches ColterBay style)
bat_theme <- theme_classic(base_size = 12) +
  theme(
    plot.background    = element_rect(fill = "white", color = NA),
    panel.background   = element_rect(fill = "white", color = NA),
    panel.grid.major.y = element_line(color = "#EBEBEB", linewidth = 0.4),
    panel.grid.major.x = element_blank(),
    axis.line          = element_line(color = "#444444", linewidth = 0.5),
    axis.ticks         = element_line(color = "#444444"),
    axis.text          = element_text(color = "#333333", size = 10),
    axis.title         = element_text(color = "#222222", size = 11),
    plot.title         = element_text(color = "#111111", size = 13,
                                      face = "bold", margin = margin(b = 3)),
    plot.subtitle      = element_text(color = "#666666", size = 9,
                                      margin = margin(b = 8)),
    legend.background  = element_rect(fill = "white", color = NA),
    legend.key         = element_rect(fill = "white", color = NA),
    legend.text        = element_text(color = "#333333", size = 10),
    legend.title       = element_text(color = "#444444", size = 9, face = "bold"),
    strip.background   = element_rect(fill = "#F2F2F2", color = NA),
    strip.text         = element_text(color = "#222222", face = "bold", size = 10),
    plot.caption       = element_text(color = "#999999", size = 8,
                                      hjust = 0, margin = margin(t = 8)),
    plot.margin        = margin(12, 14, 10, 12)
  )

# Catholic-inspired palette
col_red    <- "#7A1C2E"   # cardinal crimson
col_white  <- "#C9A84C"   # gilded gold
col_navy   <- "#1B2A4A"   # nave navy
col_purple <- "#4A1259"   # papal purple
col_ivory  <- "#F5F0E3"   # illuminated manuscript ivory

# Intensity gradient: ivory → gold → purple → crimson → navy
intensity_pal <- c("10"  = col_ivory,
                   "30"  = col_white,
                   "50"  = col_purple,
                   "70"  = col_red,
                   "100" = col_navy)
```

## Data Pipeline

```{r import, warning=FALSE, message=TRUE}
source("R/01_import_sonobat.R")
```

```{r clean, warning=FALSE, message=FALSE}
source("R/02_clean_data.R")
data <- clean_data(data_raw)
```

```{r habitat, warning=FALSE, message=FALSE}
source("R/03_extract_habitat.R")

data_env <- add_habitat(
  data,
  forest_raster = nlcd_forest,
  buffer = 50
)

data_env <- add_moonlight(data_env)
```

```{r treatment, warning=FALSE, message=FALSE}
source("R/07_add_light_treatment.R")

data_env <- data_env %>%
  left_join(
    lighting %>% select(date, color, intensity),
    by = "date"
  ) %>%
  mutate(
    color     = factor(color),
    site      = factor(site),
    intensity = factor(intensity)
  )

```

```{r distance, warning=FALSE, message=FALSE}
library(sf)

ref_point <- st_as_sf(
  data.frame(lon = -110.645134, lat = 43.904879),
  coords = c("lon", "lat"),
  crs = 4326
)

sites_sf <- st_as_sf(sites, coords = c("lon", "lat"), crs = 4326)
sites_sf <- st_transform(sites_sf, 32612)
ref_point <- st_transform(ref_point, 32612)

sites_sf$dist_m <- as.numeric(st_distance(sites_sf, ref_point))

# Pull only dist_m — avoid attaching the sf geometry class to data_env
site_dists <- as.data.frame(sites_sf) %>% dplyr::select(site, dist_m)

data_env <- data_env %>%
  left_join(site_dists, by = "site") %>%
  mutate(
    dist_km = dist_m / 1000,
    dist_cat = case_when(
      dist_km < 0.52 ~ "Close",
      dist_km < 1    ~ "Medium",
      dist_km < 2.1  ~ "Further",
      TRUE           ~ "Far"
    ),
    jd_c = jd - mean(jd, na.rm = TRUE)   # centered Julian day
  )

data_env <- data_env %>% filter(!is.na(dist_km))

data_env %>% select(site, dist_km) %>% distinct() %>% arrange(dist_km)

# Filter to focal species (matches ColterBay project)
focal_spp <- c("Laci", "Lano", "Mylu", "Epfu", "Myev", "Myvo")
data_env  <- data_env %>% filter(species %in% focal_spp)

message("Rows after species filter: ", nrow(data_env),
        "  |  Species retained: ", n_distinct(data_env$species))
```

```{r sky_brightness, warning=FALSE, message=TRUE}
source("R/09_add_sky_brightness.R")
```

```{r guild, warning=FALSE, message=FALSE}
guild_map <- tribble(
  ~species, ~guild,
  "Laci",   "Open-air",
  "Lano",   "Edge",
  "Epfu",   "Edge",
  "Mylu",   "Clutter",
  "Myev",   "Clutter",
  "Myvo",   "Clutter",
  "Myyu",   "Clutter",
  "Myci",   "Clutter"
)

data_env <- data_env %>%
  left_join(guild_map, by = "species") %>%
  mutate(guild = factor(guild, levels = c("Open-air", "Edge", "Clutter")))
```

## Models

```{r simple_model, warning=FALSE, message=FALSE}
library(glmmTMB)

simple_model1 <- glmmTMB(
  detections ~
    color * intensity +
    intensity * dist_km +
    mean_phase +
    pct_nonforest +
    brightness_dark +
    jd_c + I(jd_c^2) +
    (1 | site),
  data      = data_env,
  ziformula = ~1,
  family    = nbinom2()
)

summary(simple_model1)
```

## Community Analysis

```{r community, warning=FALSE, message=FALSE}
source("R/10_community_analysis.R")
community <- run_community_analysis(data_env, bat_theme, col_red, col_white)
```

```{r community_results, warning=FALSE, message=FALSE}
# NMDS stress
cat("NMDS stress:", round(community$nmds$stress, 3), "\n")

# PERMANOVA
community$perm

# Homogeneity of dispersion (by color)
community$disp_test

# dbRDA — global and marginal tests
community$db_global
community$db_margin

# mvabund — multivariate GLM
community$mv_anova
```

```{r community_figs, warning=FALSE, message=FALSE}
community$fig_c1   # dbRDA ordination biplot
community$fig_c2   # Species activity heatmap (color × intensity)
community$fig_c3   # Species activity heatmap (distance)
```

## Model Selection

```{r model_comparison, warning=FALSE, message=FALSE}
source("R/06_run_models.R")
models <- run_models(data_env)

AIC(models$m0, models$m1, models$m2, models$sm)

anova(models$m0, models$m1)
anova(models$m1, models$m2)
anova(models$m1, models$sm)

sim <- simulateResiduals(models$sm)
plot(sim)
testDispersion(sim)
testZeroInflation(sim)

summary(models$sm)
```

## Occurrence Thresholds

```{r thresholds, warning=FALSE, message=FALSE}
#source("R/05_data_exploration.R")
source("R/04_Occurrence_Thresholds.R")

occurrences <- run_thresholds(data_env, thresholds = c(30, 60, 120, 300))

env_data <- data_env %>%
  distinct(site, date, dist_km, mean_phase,
           color, intensity, jd_c, pct_nonforest, brightness_dark)

occurrences <- occurrences %>%
  left_join(env_data, by = c("site", "date")) %>%
  left_join(data_env %>% distinct(site, pct_forest), by = "site")

data_30  <- occurrences %>% filter(threshold_sec == 30)
data_60  <- occurrences %>% filter(threshold_sec == 60)
data_120 <- occurrences %>% filter(threshold_sec == 120)
data_300 <- occurrences %>% filter(threshold_sec == 300)
```

```{r threshold_models, warning=FALSE, message=FALSE}
threshold_formula <- occurrences ~
  color * intensity +
  intensity * dist_km +
  pct_nonforest +
  jd_c +
  I(jd_c^2) +
  mean_phase +
  brightness_dark +
  (1 | site) 

data_30_model  <- glmmTMB(threshold_formula, data = data_30, ziformula = ~1,  family = nbinom2())
data_60_model  <- glmmTMB(threshold_formula, data = data_60, ziformula = ~1,  family = nbinom2())
data_120_model <- glmmTMB(threshold_formula, data = data_120, ziformula = ~1, family = nbinom2())
data_300_model <- glmmTMB(threshold_formula, data = data_300, ziformula = ~1, family = nbinom2())

AIC(data_30_model, data_60_model, data_120_model, data_300_model)

summary(data_30_model)
```

## Diagnostics

```{r diagnostics, warning=FALSE, message=FALSE}
library(corrplot)

# Covariate correlation matrix
vars <- data_env %>%
  select(detections, pct_nonforest, jd, mean_phase, dist_km)
corrplot(cor(vars, use = "complete.obs"))

# Site-level means
site_means <- data_env %>%
  group_by(site, dist_km) %>%
  summarise(mean_det = mean(detections, na.rm = TRUE), .groups = "drop")

# Scatter with smooth
ggplot(site_means, aes(x = dist_km, y = mean_det)) +
  geom_smooth(method = "lm", se = TRUE, color = "#2E86AB",
              fill = "#DDEAF5", linewidth = 1) +
  geom_point(size = 3.5, color = "#333333") +
  ggrepel::geom_text_repel(aes(label = site), size = 3, color = "#666666") +
  labs(
    title    = "Bat Detections vs Distance from Light Source",
    subtitle = "Site-level means with linear trend",
    x        = "Distance from light (km)",
    y        = "Mean detections per night"
  ) +
  bat_theme

# Correlation test
cor.test(site_means$dist_km, site_means$mean_det, method = "spearman")


# (Distance post-hoc contrasts moved to posthoc chunk below)
```

```{r}
source("R/08b_threshold_figures.R")

fig_t1; fig_t2; fig_t3; fig_t4; fig_t5; fig_t6; fig_t7; fig_t8; fig_t9; fig_t10; fig_t11; fig_t12; fig_t13; fig_t14; fig_t15; fig_t16; fig_t17
```

## Post-hoc Contrasts

```{r posthoc, warning=FALSE, message=FALSE}
library(emmeans)

# ── 1. Color × Intensity simple effects ───────────────────────────────────────
# Red vs. White at each intensity level (the key interaction question)
em_color_int <- emmeans(simple_model1, ~ color | intensity, type = "response")

cat("=== Estimated marginal means: color at each intensity ===\n")
em_color_int

cat("\n=== Red vs. White contrast at each intensity (Holm-corrected) ===\n")
pairs(em_color_int, adjust = "holm")

# ── 2. Intensity simple effects within each color ─────────────────────────────
# Which intensity levels differ from each other, separately for red and white?
em_int_color <- emmeans(simple_model1, ~ intensity | color, type = "response")

cat("\n=== Intensity pairwise contrasts within Red (Holm-corrected) ===\n")
pairs(em_int_color, adjust = "holm")

# ── 3. Intensity × Distance interaction test ──────────────────────────────────
em_dist_trend <- emtrends(simple_model1, ~ intensity,
                           var = "dist_km", type = "response")

cat("\n=== Distance slope (km) by intensity level ===\n")
em_dist_trend

cat("\n=== Pairwise intensity distance-slope contrasts (Holm-corrected) ===\n")
pairs(em_dist_trend, adjust = "holm")

# Contrasts at meaningful distances
em_dist_points <- emmeans(
  simple_model1,
  ~ intensity | dist_km,
  at   = list(dist_km = c(0.3, 0.7, 2.0)),
  type = "response"
)

cat("\n=== Intensity contrasts at close (0.3 km), medium (0.7 km), far (2.0 km) ===\n")
pairs(em_dist_points, adjust = "holm")
```

## Distance × Intensity Figure

```{r dist_intensity_fig, warning=FALSE, message=FALSE}
library(emmeans)

# Predicted detections across the observed distance gradient, by intensity
dist_range <- range(data_env$dist_km, na.rm = TRUE)
dist_grid  <- seq(dist_range[1], dist_range[2], length.out = 60)

em_dist_int <- emmeans(
  simple_model1,
  ~ intensity | dist_km,
  at      = list(dist_km = dist_grid),
  type    = "response"
)

pred_df <- as.data.frame(em_dist_int) %>%
  rename(lower_CL = asymp.LCL, upper_CL = asymp.UCL)

ggplot(pred_df, aes(x = dist_km, y = response,
                    color = factor(intensity), fill = factor(intensity))) +
  geom_ribbon(aes(ymin = lower_CL, ymax = upper_CL), alpha = 0.12, color = NA) +
  geom_line(linewidth = 1.1) +
  scale_color_manual(values = intensity_pal, name = "Intensity (%)") +
  scale_fill_manual(values  = intensity_pal, name = "Intensity (%)") +
  scale_x_continuous(breaks = scales::pretty_breaks(n = 5)) +
  labs(
    title    = "Predicted Bat Detections by Distance and Light Intensity",
    subtitle = "Marginal means ± 95% CI from negative binomial GLMM",
    x        = "Distance from Colter Bay parking lot (km)",
    y        = "Predicted detections per night",
    caption  = "Other covariates held at their means"
  ) +
  bat_theme +
  theme(legend.position = "right")

ggsave("output/figures/fig9_dist_intensity.png",
       last_plot(), width = 8, height = 5, dpi = 300, bg = "white")
```

## Guild Analysis

```{r guild_model, warning=FALSE, message=FALSE}
guild_model <- glmmTMB(
  detections ~
    guild * intensity +
    guild * dist_km +
    mean_phase +
    pct_nonforest +
    jd_c + I(jd_c^2) +
    (1 | site),
  data      = data_env,
  ziformula = ~1,
  family    = nbinom2()
)

summary(guild_model)
```

```{r guild_posthoc, warning=FALSE, message=FALSE}
# Intensity effect by guild
em_guild_intensity <- emmeans(guild_model, ~ intensity | guild, type = "response")
cat("=== Intensity effect within each guild ===\n")
em_guild_intensity
pairs(em_guild_intensity, adjust = "holm")

# Distance slope by guild
em_guild_dist <- emtrends(guild_model, ~ guild, var = "dist_km", type = "response")
cat("\n=== Distance slope by guild ===\n")
em_guild_dist
pairs(em_guild_dist, adjust = "holm")
```

```{r guild_figs, warning=FALSE, message=FALSE}
# Fig A: intensity effect by guild
em_guild_intensity_df <- as.data.frame(em_guild_intensity) %>%
  rename(lower_CL = asymp.LCL, upper_CL = asymp.UCL) %>%
  mutate(intensity = factor(intensity, levels = c("10","30","50","70","100")))

fig_guild_intensity <- ggplot(em_guild_intensity_df,
                              aes(x = guild, y = response,
                                  color = intensity, group = intensity)) +
  geom_line(linewidth = 0.8, position = position_dodge(0.4)) +
  geom_point(size = 3.5, position = position_dodge(0.4)) +
  geom_errorbar(aes(ymin = lower_CL, ymax = upper_CL),
                width = 0.15, linewidth = 0.7,
                position = position_dodge(0.4)) +
  scale_color_manual(values = intensity_pal, name = "Intensity (%)") +
  labs(
    title    = "Light Intensity Effect by Foraging Guild",
    subtitle = "Marginal means ± 95% CI from guild × intensity model",
    x        = "Foraging guild",
    y        = "Predicted detections per night"
  ) +
  bat_theme +
  theme(legend.position = "top")

# Fig B: distance trend by guild
dist_grid_guild <- seq(min(data_env$dist_km, na.rm = TRUE),
                       max(data_env$dist_km, na.rm = TRUE),
                       length.out = 60)

em_guild_dist_df <- emmeans(guild_model, ~ guild + dist_km,
                             at   = list(dist_km = dist_grid_guild),
                             type = "response") %>%
  as.data.frame() %>%
  rename(lower_CL = asymp.LCL, upper_CL = asymp.UCL)

fig_guild_dist <- ggplot(em_guild_dist_df,
                         aes(x = dist_km, y = response,
                             color = guild, fill = guild)) +
  geom_ribbon(aes(ymin = lower_CL, ymax = upper_CL),
              alpha = 0.15, color = NA) +
  geom_line(linewidth = 1.1) +
  scale_color_manual(values = c("Open-air" = col_navy,
                                 "Edge"     = col_red,
                                 "Clutter"  = col_purple),
                     name = "Guild") +
  scale_fill_manual(values  = c("Open-air" = col_navy,
                                 "Edge"     = col_red,
                                 "Clutter"  = col_purple),
                    guide   = "none") +
  labs(
    title    = "Bat Activity Across Distance Gradient by Foraging Guild",
    subtitle = "Marginal predictions ± 95% CI · averaged over light color",
    x        = "Distance from light source (km)",
    y        = "Predicted detections per night"
  ) +
  bat_theme +
  theme(legend.position = "top")

fig_guild_intensity
fig_guild_dist

ggsave("output/figures/fig_guild_intensity.png",
       fig_guild_intensity, width = 7, height = 5, dpi = 300, bg = "white")
ggsave("output/figures/fig_guild_dist.png",
       fig_guild_dist,      width = 8, height = 5, dpi = 300, bg = "white")
```

## Species Models

```{r species_models, warning=FALSE, message=FALSE}
source("R/11_species_models.R")
sp_results <- run_species_models(data_env, bat_theme, col_red, col_white)
```

```{r species_figs, warning=FALSE, message=FALSE}
sp_results$fig_s1   # Color effect by species
sp_results$fig_s2   # Intensity slopes by species
sp_results$fig_s3   # White:red ratio heatmap
sp_results$fig_s4a  # Predicted detections ~ dist_km × color × intensity (species facets)
sp_results$fig_s4b  # Same, species × color grid
```

## Optimal Light Treatment by Site

```{r optimal_treatment, warning=FALSE, message=FALSE}
library(emmeans)

site_dists <- data_env %>%
  distinct(site, dist_km) %>%
  arrange(dist_km)

# Predict detections at every color × intensity combo at each site's dist_km
em_opt <- emmeans(
  simple_model1,
  ~ color * intensity + dist_km,
  at   = list(dist_km = site_dists$dist_km),
  type = "response"
) %>%
  as.data.frame() %>%
  rename(lower_CL = asymp.LCL, upper_CL = asymp.UCL) %>%
  left_join(site_dists, by = "dist_km") %>%
  mutate(
    combo     = paste0(ifelse(color == "R", "Red", "White"), " / ", intensity, "%"),
    site      = factor(site, levels = site_dists$site),
    intensity = factor(intensity, levels = c("10","30","50","70","100"))
  )

# Flag the best combo per site
best_combo <- em_opt %>%
  group_by(site) %>%
  slice_max(response, n = 1) %>%
  ungroup() %>%
  select(site, combo, response, lower_CL, upper_CL)

cat("=== Best predicted light treatment per site ===\n")
print(best_combo)

# Heatmap: site (y) × combo (x), fill = predicted detections
combo_order <- em_opt %>%
  group_by(combo) %>%
  summarise(mean_r = mean(response)) %>%
  arrange(mean_r) %>%
  pull(combo)

em_opt <- em_opt %>%
  mutate(
    combo          = factor(combo, levels = combo_order),
    response_scaled = (response - min(response)) / (max(response) - min(response))
  )

best_marks <- em_opt %>%
  group_by(site) %>%
  slice_max(response, n = 1) %>%
  ungroup()

ggplot(em_opt, aes(x = combo, y = fct_rev(site), fill = response_scaled)) +
  geom_tile(color = "white", linewidth = 0.4) +
  geom_text(data = best_marks, label = "*", color = "white",
            size = 4, fontface = "bold") +
  scale_fill_gradient(low = col_ivory, high = col_navy,
                      name = "Relative\nactivity") +
  labs(
    title    = "Predicted Bat Detections by Light Treatment and Site",
    subtitle = "* = best treatment per site · sites ordered by increasing distance",
    x        = "Light color / intensity",
    y        = NULL
  ) +
  bat_theme +
  theme(
    axis.text.x  = element_text(angle = 45, hjust = 1, size = 9),
    panel.grid   = element_blank(),
    axis.line    = element_blank(),
    axis.ticks   = element_blank()
  )

ggsave("output/figures/fig_optimal_treatment.png",
       last_plot(), width = 10, height = 5, dpi = 300, bg = "white")
```

```{r optimal_by_zone, warning=FALSE, message=FALSE}
zone_sites <- list(
  Close  = c("GRTE17","GRTE16","GRTE05","GRTE01","GRTE06","GRTE02"),
  Medium = c("GRTE08","GRTE03","GRTE09","GRTE11","GRTE10"),
  Far    = c("GRTE15","GRTE14","GRTE13")
)

zone_lookup <- stack(zone_sites) %>%
  rename(site = values, zone = ind) %>%
  mutate(zone = factor(zone, levels = c("Close","Medium","Far")))

# Average predicted detections across sites within each zone
em_zone <- em_opt %>%
  left_join(zone_lookup, by = "site") %>%
  group_by(zone, color, intensity, combo) %>%
  summarise(
    response  = mean(response,  na.rm = TRUE),
    lower_CL  = mean(lower_CL,  na.rm = TRUE),
    upper_CL  = mean(upper_CL,  na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(combo = factor(combo, levels = combo_order))

best_zone <- em_zone %>%
  group_by(zone) %>%
  slice_max(response, n = 1) %>%
  ungroup()

cat("=== Best predicted light treatment per distance zone ===\n")
best_zone %>% select(zone, combo, response, lower_CL, upper_CL) %>% print()

ggplot(em_zone, aes(x = response, y = combo, color = combo)) +
  geom_linerange(aes(xmin = lower_CL, xmax = upper_CL),
                 linewidth = 0.6, alpha = 0.6) +
  geom_point(size = 3) +
  geom_point(data = best_zone, size = 5, shape = 21,
             fill = NA, stroke = 1.3, color = "#111111") +
  scale_color_manual(values = treatment_pal, guide = "none") +
  facet_wrap(~ zone, nrow = 1, scales = "free_x") +
  labs(
    title    = "Optimal Light Treatment by Distance Zone",
    subtitle = "Circle outline = best treatment · bars = 95% CI · averaged across sites within zone",
    x        = "Predicted bat passes per night",
    y        = NULL
  ) +
  bat_theme +
  theme(strip.text = element_text(size = 10, face = "bold"))

ggsave("output/figures/fig_optimal_treatment_by_zone.png",
       last_plot(), width = 10, height = 6, dpi = 300, bg = "white")
```

```{r close_vs_far, warning=FALSE, message=FALSE}
# Define distance zones by site groupings
zone_sites <- list(
  Close  = c("GRTE17","GRTE16","GRTE05","GRTE01","GRTE06","GRTE02"),
  Medium = c("GRTE08","GRTE03","GRTE09","GRTE11","GRTE10"),
  Far    = c("GRTE15","GRTE14","GRTE13")
)

zone_km <- map_dbl(zone_sites, ~ mean(site_dists$dist_km[site_dists$site %in% .x], na.rm = TRUE))
cat("Mean dist_km per zone:\n"); print(round(zone_km, 2))

df_close_far <- map2_dfr(sp_results$sp_models, names(sp_results$sp_models),
  function(mod, sp) {
    tryCatch({
      emmeans(mod, ~ color * intensity + dist_km,
              at   = list(dist_km = unname(zone_km)),
              type = "response") %>%
        as.data.frame() %>%
        rename(lower_CL = asymp.LCL, upper_CL = asymp.UCL) %>%
        mutate(
          species       = sp,
          distance_zone = case_when(
            dist_km == zone_km["Close"]  ~ paste0("Close (", round(zone_km["Close"],  2), " km)"),
            dist_km == zone_km["Medium"] ~ paste0("Medium (", round(zone_km["Medium"], 2), " km)"),
            dist_km == zone_km["Far"]    ~ paste0("Far (", round(zone_km["Far"],    2), " km)")
          ),
          distance_zone = factor(distance_zone,
                                 levels = c(paste0("Close (", round(zone_km["Close"], 2), " km)"),
                                            paste0("Medium (", round(zone_km["Medium"], 2), " km)"),
                                            paste0("Far (", round(zone_km["Far"], 2), " km)"))),
          combo = paste0(ifelse(color == "R", "Red", "White"), " / ", intensity, "%"),
          combo = factor(combo, levels = combo_order)
        )
    }, error = function(e) {
      cat("Skipping", sp, ":", conditionMessage(e), "\n"); NULL
    })
  })

# Best combo per species × zone for annotation
best_close_far <- df_close_far %>%
  group_by(species, distance_zone) %>%
  slice_max(response, n = 1) %>%
  ungroup()

cat("=== Best treatment per species at close vs. far distance ===\n")
best_close_far %>%
  select(species, distance_zone, combo, response, lower_CL, upper_CL) %>%
  arrange(species, distance_zone) %>%
  print(n = Inf)

treatment_pal <- c(
  "Red / 10%"    = "#7A1C2E", "Red / 30%"    = "#A84050",
  "Red / 50%"    = "#C96070", "Red / 70%"    = "#E08090",
  "Red / 100%"   = "#F0B0BA", "White / 10%"  = "#F5F0E3",
  "White / 30%"  = "#C9A84C", "White / 50%"  = "#A08030",
  "White / 70%"  = "#706020", "White / 100%" = "#4A1259"
)

sp_zone_figs <- map(unique(df_close_far$species), function(sp) {
  dat      <- filter(df_close_far, species == sp)
  dat_best <- filter(best_close_far, species == sp)

  ggplot(dat, aes(x = response, y = combo, color = combo)) +
    geom_linerange(aes(xmin = lower_CL, xmax = upper_CL),
                   linewidth = 0.6, alpha = 0.6) +
    geom_point(size = 3) +
    geom_point(data = dat_best, size = 5, shape = 21,
               fill = NA, stroke = 1.3, color = "#111111") +
    scale_color_manual(values = treatment_pal, guide = "none") +
    facet_wrap(~ distance_zone, nrow = 1, scales = "free_x") +
    labs(
      title    = sp,
      subtitle = "Circle outline = best treatment · bars = 95% CI",
      x        = "Predicted detections per night",
      y        = NULL
    ) +
    bat_theme +
    theme(strip.text = element_text(size = 9, face = "bold"))
})
names(sp_zone_figs) <- unique(df_close_far$species)

# Display each figure
walk(sp_zone_figs, print)

# Save individually
dir.create("output/figures/species_zone", recursive = TRUE, showWarnings = FALSE)
iwalk(sp_zone_figs, ~ ggsave(
  paste0("output/figures/species_zone/", .y, "_zone_treatment.png"),
  .x, width = 10, height = 5, dpi = 300, bg = "white"
))
```

```{r optimal_by_species, warning=FALSE, message=FALSE}
# Same heatmap per species using the per-species GLMMs
em_sp_opt <- map2_dfr(sp_results$sp_models, names(sp_results$sp_models),
  function(mod, sp) {
    tryCatch({
      emmeans(mod, ~ color * intensity + dist_km,
              at   = list(dist_km = site_dists$dist_km),
              type = "response") %>%
        as.data.frame() %>%
        rename(lower_CL = asymp.LCL, upper_CL = asymp.UCL) %>%
        left_join(site_dists, by = "dist_km") %>%
        mutate(species = sp)
    }, error = function(e) {
      cat("Skipping", sp, ":", conditionMessage(e), "\n"); NULL
    })
  }) %>%
  mutate(
    combo = paste0(ifelse(color == "R", "Red", "White"), " / ", intensity, "%"),
    combo = factor(combo, levels = combo_order),
    site  = factor(site, levels = site_dists$site)
  ) %>%
  group_by(species) %>%
  mutate(response_scaled = (response - min(response)) / (max(response) - min(response))) %>%
  ungroup()

best_sp <- em_sp_opt %>%
  group_by(species, site) %>%
  slice_max(response, n = 1) %>%
  ungroup()

ggplot(em_sp_opt, aes(x = combo, y = fct_rev(site), fill = response_scaled)) +
  geom_tile(color = "white", linewidth = 0.3) +
  geom_text(data = best_sp, label = "*", color = "white",
            size = 3, fontface = "bold") +
  scale_fill_gradient(low = col_ivory, high = col_navy,
                      name = "Relative\nactivity") +
  facet_wrap(~ species, nrow = 2) +
  labs(
    title    = "Predicted Detections by Light Treatment and Site — Per Species",
    subtitle = "* = best treatment per site · sites ordered by increasing distance",
    x        = "Light color / intensity",
    y        = NULL
  ) +
  bat_theme +
  theme(
    axis.text.x     = element_text(angle = 45, hjust = 1, size = 7),
    strip.text      = element_text(size = 9, face = "bold"),
    panel.grid      = element_blank(),
    axis.line       = element_blank(),
    axis.ticks      = element_blank(),
    legend.position = "right"
  )

ggsave("output/figures/fig_optimal_treatment_by_species.png",
       last_plot(), width = 16, height = 8, dpi = 300, bg = "white")
```

## Figures

```{r figures, message=FALSE, warning=FALSE}
source("R/08_figures.R")

fig1; fig2; fig3; fig3b; fig4; fig5; fig6; fig7; fig8; fig9; fig10; fig11; fig12; fig13; fig14; fig15; fig16; fig17; fig18; fig19; fig20; fig21; fig22; fig23; fig24; fig25; fig26
```

## Species Key

| Code | Species                                          |
|------|--------------------------------------------------|
| Myyu | Yuma myotis (possibly Myvo)                      |
| Myvo | *Myotis volans* \*                               |
| Myth | *Myotis thysanodes* (possibly Myev)              |
| Mylu | *Myotis lucifugus* \*                            |
| Myev | *Myotis evotis* \*                               |
| Myci | *Myotis ciliolabrum*                             |
| Myca | *Myotis californicus*                            |
| Lano | *Lasionycteris noctivagans*                      |
| Laci | *Lasiurus cinereus* \*                           |
| Euma | *Euderma maculatum*                              |
| Epfu | *Eptesicus fuscus* \*                            |
| Anpa | *Antrozous pallidus* (possibly *L. noctivagans*) |

------------------------------------------------------------------------
