---
title: "Distance Project"
author: "Nigel K Anderson"
format: html
editor: visual
---

```{r setup, include=FALSE}
library(tidyverse)
library(ggplot2)

# Shared theme + palette (matches ColterBay style)
bat_theme <- theme_classic(base_size = 12) +
  theme(
    plot.background    = element_rect(fill = "white", color = NA),
    panel.background   = element_rect(fill = "white", color = NA),
    panel.grid.major.y = element_line(color = "#EBEBEB", linewidth = 0.4),
    panel.grid.major.x = element_blank(),
    axis.line          = element_line(color = "#444444", linewidth = 0.5),
    axis.ticks         = element_line(color = "#444444"),
    axis.text          = element_text(color = "#333333", size = 10),
    axis.title         = element_text(color = "#222222", size = 11),
    plot.title         = element_text(color = "#111111", size = 13,
                                      face = "bold", margin = margin(b = 3)),
    plot.subtitle      = element_text(color = "#666666", size = 9,
                                      margin = margin(b = 8)),
    legend.background  = element_rect(fill = "white", color = NA),
    legend.key         = element_rect(fill = "white", color = NA),
    legend.text        = element_text(color = "#333333", size = 10),
    legend.title       = element_text(color = "#444444", size = 9, face = "bold"),
    strip.background   = element_rect(fill = "#F2F2F2", color = NA),
    strip.text         = element_text(color = "#222222", face = "bold", size = 10),
    plot.caption       = element_text(color = "#999999", size = 8,
                                      hjust = 0, margin = margin(t = 8)),
    plot.margin        = margin(12, 14, 10, 12)
  )

col_red   <- "#C0392B"
col_white <- "#2E86AB"
```

## Data Pipeline

```{r import, warning=FALSE, message=TRUE}
source("R/01_import_sonobat.R")
```

```{r clean, warning=FALSE, message=FALSE}
source("R/02_clean_data.R")
data <- clean_data(data_raw)
```

```{r habitat, warning=FALSE, message=FALSE}
source("R/03_extract_habitat.R")

data_env <- add_habitat(
  data,
  forest_raster = nlcd_forest,
  buffer = 50
)

data_env <- add_moonlight(data_env)
```

```{r treatment, warning=FALSE, message=FALSE}
source("R/07_add_light_treatment.R")

data_full <- readRDS("data/data_full.rds")

data_env <- data_env %>%
  left_join(
    lighting %>% select(date, color, intensity),
    by = "date"
  ) %>%
  mutate(
    color     = factor(color),
    site      = factor(site),
    intensity = factor(intensity)
  )
```

```{r distance, warning=FALSE, message=FALSE}
library(sf)

ref_point <- st_as_sf(
  data.frame(lon = -110.645134, lat = 43.904879),
  coords = c("lon", "lat"),
  crs = 4326
)

sites_sf <- st_as_sf(sites, coords = c("lon", "lat"), crs = 4326)
sites_sf <- st_transform(sites_sf, 32612)
ref_point <- st_transform(ref_point, 32612)

sites_sf$dist_m <- as.numeric(st_distance(sites_sf, ref_point))

data_env <- data_env %>%
  left_join(sites_sf, by = "site") %>%
  mutate(
    dist_km = dist_m / 1000,
    dist_cat = case_when(
      dist_km < 0.52 ~ "Close",
      dist_km < 1    ~ "Medium",
      dist_km < 2.1  ~ "Further",
      TRUE           ~ "Far"
    ),
    jd_c = jd - mean(jd, na.rm = TRUE)   # centered Julian day
  )

data_env %>% select(site, dist_km) %>% distinct() %>% arrange(dist_km)
```

## Models

```{r simple_model, warning=FALSE, message=FALSE}
library(glmmTMB)

simple_model1 <- glmmTMB(
  detections ~
    color * intensity +
    mean_phase +
    pct_nonforest +
    jd_c + I(jd_c^2) +
    dist_cat +
    (1 | site),
  data      = data_env,
  ziformula = ~1,
  family    = nbinom2()
)

summary(simple_model1)
```

## Community Analysis

```{r community, warning=FALSE, message=FALSE}
library(vegan)
library(indicspecies)

data_env$intensity <- as.numeric(as.character(data_env$intensity))

comm_df <- data_env %>%
  group_by(site, color, intensity, species) %>%
  summarise(n = n(), .groups = "drop")

comm_wide <- comm_df %>%
  tidyr::pivot_wider(
    names_from  = species,
    values_from = n,
    values_fill = 0
  )

meta        <- comm_wide %>% select(site, color, intensity)
comm_matrix <- comm_wide %>% select(-site, -color, intensity)

set.seed(123)
nmds <- metaMDS(comm_matrix, distance = "bray", k = 2, trymax = 100)
nmds$stress

# NMDS biplot
nmds_sites          <- as.data.frame(scores(nmds, display = "sites"))
nmds_sites          <- cbind(nmds_sites, meta)
nmds_spp            <- as.data.frame(scores(nmds, display = "species"))
nmds_spp$species    <- rownames(nmds_spp)

ggplot(nmds_sites, aes(NMDS1, NMDS2)) +
  geom_point(aes(color = color), size = 3) +
  geom_segment(
    data        = nmds_spp,
    aes(x = 0, xend = NMDS1, y = 0, yend = NMDS2),
    arrow       = arrow(length = unit(0.2, "cm")),
    inherit.aes = FALSE
  ) +
  geom_text(
    data        = nmds_spp,
    aes(x = NMDS1, y = NMDS2, label = species),
    size        = 3,
    inherit.aes = FALSE
  ) +
  scale_color_manual(
    values = c(R = col_red, W = col_white),
    labels = c(R = "Red", W = "White"),
    name   = "Light color"
  ) +
  labs(title = "NMDS: Bat Community Composition") +
  bat_theme

# Permanova
adonis2(comm_matrix ~ color * intensity,        data = meta, method = "bray")
adonis2(comm_matrix ~ color * intensity + site, data = meta, method = "bray")

# Dispersion
dist_mat <- vegdist(comm_matrix, method = "bray")
bd       <- betadisper(dist_mat, meta$color)
anova(bd)

# Indicator species
indval <- multipatt(comm_matrix, meta$color, control = how(nperm = 999))
summary(indval)
```

## Model Selection

```{r model_comparison, warning=FALSE, message=FALSE}
source("R/06_run_models.R")
models <- run_models(data_env)

AIC(models$m0, models$m1, models$m2, models$sm)

anova(models$m0, models$m1)
anova(models$m1, models$m2)
anova(models$m1, models$sm)

sim <- simulateResiduals(models$sm)
plot(sim)
testDispersion(sim)
testZeroInflation(sim)

summary(models$sm)
```

## Occurrence Thresholds

```{r thresholds, warning=FALSE, message=FALSE}
source("R/05_data_exploration.R")
source("R/04_Occurrence_Thresholds.R")

occurrences <- run_thresholds(data_env, thresholds = c(30, 60, 120, 300))

env_data <- data_env %>%
  distinct(site, date, mean_moonlight, mean_phase,
           color, intensity, year, jd, pct_forest)

occurrences <- occurrences %>%
  left_join(env_data, by = c("site", "date")) %>%
  left_join(data_env %>% distinct(site, pct_forest), by = "site")

data_30  <- occurrences %>% filter(threshold_sec == 30)
data_60  <- occurrences %>% filter(threshold_sec == 60)
data_120 <- occurrences %>% filter(threshold_sec == 120)
data_300 <- occurrences %>% filter(threshold_sec == 300)
```

```{r threshold_models, warning=FALSE, message=FALSE}
threshold_formula <- occurrences ~
  color * intensity * mean_phase +
  pct_forest +
  jd +
  jd^2 +
  mean_moonlight +
  (1 | site) +
  (1 | year)

data_30_model  <- glmmTMB(threshold_formula, data = data_30,  family = nbinom2())
data_60_model  <- glmmTMB(threshold_formula, data = data_60,  family = nbinom2())
data_120_model <- glmmTMB(threshold_formula, data = data_120, family = nbinom2())
data_300_model <- glmmTMB(threshold_formula, data = data_300, family = nbinom2())

AIC(data_30_model, data_60_model, data_120_model, data_300_model)

summary(data_30_model)
```

## Diagnostics

```{r diagnostics, warning=FALSE, message=FALSE}
library(corrplot)
library(emmeans)

# Covariate correlation matrix
vars <- data_env %>%
  select(detections, pct_nonforest, jd, mean_phase, dist_km)
corrplot(cor(vars, use = "complete.obs"))

# Site-level means
site_means <- data_env %>%
  group_by(site, dist_km) %>%
  summarise(mean_det = mean(detections, na.rm = TRUE), .groups = "drop")

# Scatter with smooth
ggplot(site_means, aes(x = dist_km, y = mean_det)) +
  geom_smooth(method = "lm", se = TRUE, color = "#2E86AB",
              fill = "#DDEAF5", linewidth = 1) +
  geom_point(size = 3.5, color = "#333333") +
  ggrepel::geom_text_repel(aes(label = site), size = 3, color = "#666666") +
  labs(
    title    = "Bat Detections vs Distance from Light Source",
    subtitle = "Site-level means with linear trend",
    x        = "Distance from light (km)",
    y        = "Mean detections per night"
  ) +
  bat_theme

# Correlation test
cor.test(site_means$dist_km, site_means$mean_det, method = "spearman")


# Distance post-hoc contrasts
emmeans(simple_model1, ~dist_cat)
pairs(emmeans(simple_model1, ~dist_cat))
```

## Figures

```{r figures, message=FALSE, warning=FALSE}
source("R/08_figures.R")

fig1; fig2; fig3; fig4; fig5; fig6; fig7; fig8
```

## Species Key

| Code | Species                                          |
|------|--------------------------------------------------|
| Myyu | Yuma myotis (possibly Myvo)                      |
| Myvo | *Myotis volans* \*                               |
| Myth | *Myotis thysanodes* (possibly Myev)              |
| Mylu | *Myotis lucifugus* \*                            |
| Myev | *Myotis evotis* \*                               |
| Myci | *Myotis ciliolabrum*                             |
| Myca | *Myotis californicus*                            |
| Lano | *Euderma maculatum*                              |
| Laci | *Lasiurus cinereus* \*                           |
| Euma | *Euderma maculatum*                              |
| Epfu | *Eptesicus fuscus* \*                            |
| Anpa | *Antrozous pallidus* (possibly *L. noctivagans*) |

------------------------------------------------------------------------

*At the end of every day:*

```         
git add .
git commit -m "updated detection pipeline"
git push
```
