---
title: "Distance Project"
author: "Nigel K Anderson"
format: html
editor: visual
---

```{r setup, include=FALSE}
library(tidyverse)
library(ggplot2)

# Shared theme + palette (matches ColterBay style)
bat_theme <- theme_classic(base_size = 12) +
  theme(
    plot.background    = element_rect(fill = "white", color = NA),
    panel.background   = element_rect(fill = "white", color = NA),
    panel.grid.major.y = element_line(color = "#EBEBEB", linewidth = 0.4),
    panel.grid.major.x = element_blank(),
    axis.line          = element_line(color = "#444444", linewidth = 0.5),
    axis.ticks         = element_line(color = "#444444"),
    axis.text          = element_text(color = "#333333", size = 10),
    axis.title         = element_text(color = "#222222", size = 11),
    plot.title         = element_text(color = "#111111", size = 13,
                                      face = "bold", margin = margin(b = 3)),
    plot.subtitle      = element_text(color = "#666666", size = 9,
                                      margin = margin(b = 8)),
    legend.background  = element_rect(fill = "white", color = NA),
    legend.key         = element_rect(fill = "white", color = NA),
    legend.text        = element_text(color = "#333333", size = 10),
    legend.title       = element_text(color = "#444444", size = 9, face = "bold"),
    strip.background   = element_rect(fill = "#F2F2F2", color = NA),
    strip.text         = element_text(color = "#222222", face = "bold", size = 10),
    plot.caption       = element_text(color = "#999999", size = 8,
                                      hjust = 0, margin = margin(t = 8)),
    plot.margin        = margin(12, 14, 10, 12)
  )

col_red   <- "#C0392B"
col_white <- "#2E86AB"
```

## Data Pipeline

```{r import, warning=FALSE, message=TRUE}
source("R/01_import_sonobat.R")
```

```{r clean, warning=FALSE, message=FALSE}
source("R/02_clean_data.R")
data <- clean_data(data_raw)
```

```{r habitat, warning=FALSE, message=FALSE}
source("R/03_extract_habitat.R")

data_env <- add_habitat(
  data,
  forest_raster = nlcd_forest,
  buffer = 50
)

data_env <- add_moonlight(data_env)
```

```{r treatment, warning=FALSE, message=FALSE}
source("R/07_add_light_treatment.R")

data_full <- readRDS("data/data_full.rds")

data_env <- data_env %>%
  left_join(
    lighting %>% select(date, color, intensity),
    by = "date"
  ) %>%
  mutate(
    color     = factor(color),
    site      = factor(site),
    intensity = factor(intensity)
  )
```

```{r distance, warning=FALSE, message=FALSE}
library(sf)

ref_point <- st_as_sf(
  data.frame(lon = -110.645134, lat = 43.904879),
  coords = c("lon", "lat"),
  crs = 4326
)

sites_sf <- st_as_sf(sites, coords = c("lon", "lat"), crs = 4326)
sites_sf <- st_transform(sites_sf, 32612)
ref_point <- st_transform(ref_point, 32612)

sites_sf$dist_m <- as.numeric(st_distance(sites_sf, ref_point))

# Pull only dist_m — avoid attaching the sf geometry class to data_env
site_dists <- as.data.frame(sites_sf) %>% dplyr::select(site, dist_m)

data_env <- data_env %>%
  left_join(site_dists, by = "site") %>%
  mutate(
    dist_km = dist_m / 1000,
    dist_cat = case_when(
      dist_km < 0.52 ~ "Close",
      dist_km < 1    ~ "Medium",
      dist_km < 2.1  ~ "Further",
      TRUE           ~ "Far"
    ),
    jd_c = jd - mean(jd, na.rm = TRUE)   # centered Julian day
  )

data_env %>% select(site, dist_km) %>% distinct() %>% arrange(dist_km)

# Filter to focal species (matches ColterBay project)
focal_spp <- c("Laci", "Lano", "Mylu", "Epfu", "Myev", "Myvo", "Myyu", "Myci")
data_env  <- data_env %>% filter(species %in% focal_spp)

message("Rows after species filter: ", nrow(data_env),
        "  |  Species retained: ", n_distinct(data_env$species))
```

```{r sky_brightness, warning=FALSE, message=TRUE}
source("R/09_add_sky_brightness.R")
```

## Models

```{r simple_model, warning=FALSE, message=FALSE}
library(glmmTMB)

simple_model1 <- glmmTMB(
  detections ~
    color * intensity +
    mean_phase +
    pct_nonforest +
    brightness_dark +
    jd_c + I(jd_c^2) +
    dist_cat +
    (1 | site),
  data      = data_env,
  ziformula = ~1,
  family    = nbinom2()
)

summary(simple_model1)
```

## Community Analysis

```{r community, warning=FALSE, message=FALSE}
source("R/10_community_analysis.R")
community <- run_community_analysis(data_env, bat_theme, col_red, col_white)
```

```{r community_results, warning=FALSE, message=FALSE}
# NMDS stress
cat("NMDS stress:", round(community$nmds$stress, 3), "\n")

# PERMANOVA
community$perm

# Homogeneity of dispersion (by color)
community$disp_test

# dbRDA — global and marginal tests
community$db_global
community$db_margin

# mvabund — multivariate GLM
community$mv_anova
```

```{r community_figs, warning=FALSE, message=FALSE}
community$fig_c1   # dbRDA ordination biplot
community$fig_c2   # Species activity heatmap (color × intensity)
community$fig_c3   # Species activity heatmap (distance)
```

## Model Selection

```{r model_comparison, warning=FALSE, message=FALSE}
source("R/06_run_models.R")
models <- run_models(data_env)

AIC(models$m0, models$m1, models$m2, models$sm)

anova(models$m0, models$m1)
anova(models$m1, models$m2)
anova(models$m1, models$sm)

sim <- simulateResiduals(models$sm)
plot(sim)
testDispersion(sim)
testZeroInflation(sim)

summary(models$sm)
```

## Occurrence Thresholds

```{r thresholds, warning=FALSE, message=FALSE}
source("R/05_data_exploration.R")
source("R/04_Occurrence_Thresholds.R")

occurrences <- run_thresholds(data_env, thresholds = c(30, 60, 120, 300))

env_data <- data_env %>%
  distinct(site, date, mean_moonlight, mean_phase,
           color, intensity, year, jd, pct_forest)

occurrences <- occurrences %>%
  left_join(env_data, by = c("site", "date")) %>%
  left_join(data_env %>% distinct(site, pct_forest), by = "site")

data_30  <- occurrences %>% filter(threshold_sec == 30)
data_60  <- occurrences %>% filter(threshold_sec == 60)
data_120 <- occurrences %>% filter(threshold_sec == 120)
data_300 <- occurrences %>% filter(threshold_sec == 300)
```

```{r threshold_models, warning=FALSE, message=FALSE}
threshold_formula <- occurrences ~
  color * intensity * mean_phase +
  pct_forest +
  jd +
  jd^2 +
  mean_moonlight +
  (1 | site) +
  (1 | year)

data_30_model  <- glmmTMB(threshold_formula, data = data_30,  family = nbinom2())
data_60_model  <- glmmTMB(threshold_formula, data = data_60,  family = nbinom2())
data_120_model <- glmmTMB(threshold_formula, data = data_120, family = nbinom2())
data_300_model <- glmmTMB(threshold_formula, data = data_300, family = nbinom2())

AIC(data_30_model, data_60_model, data_120_model, data_300_model)

summary(data_30_model)
```

## Diagnostics

```{r diagnostics, warning=FALSE, message=FALSE}
library(corrplot)
library(emmeans)

# Covariate correlation matrix
vars <- data_env %>%
  select(detections, pct_nonforest, jd, mean_phase, dist_km)
corrplot(cor(vars, use = "complete.obs"))

# Site-level means
site_means <- data_env %>%
  group_by(site, dist_km) %>%
  summarise(mean_det = mean(detections, na.rm = TRUE), .groups = "drop")

# Scatter with smooth
ggplot(site_means, aes(x = dist_km, y = mean_det)) +
  geom_smooth(method = "lm", se = TRUE, color = "#2E86AB",
              fill = "#DDEAF5", linewidth = 1) +
  geom_point(size = 3.5, color = "#333333") +
  ggrepel::geom_text_repel(aes(label = site), size = 3, color = "#666666") +
  labs(
    title    = "Bat Detections vs Distance from Light Source",
    subtitle = "Site-level means with linear trend",
    x        = "Distance from light (km)",
    y        = "Mean detections per night"
  ) +
  bat_theme

# Correlation test
cor.test(site_means$dist_km, site_means$mean_det, method = "spearman")


# (Distance post-hoc contrasts moved to posthoc chunk below)
```

## Post-hoc Contrasts

```{r posthoc, warning=FALSE, message=FALSE}
library(emmeans)

# ── 1. Color × Intensity simple effects ───────────────────────────────────────
# Red vs. White at each intensity level (the key interaction question)
em_color_int <- emmeans(simple_model1, ~ color | intensity, type = "response")

cat("=== Estimated marginal means: color at each intensity ===\n")
em_color_int

cat("\n=== Red vs. White contrast at each intensity (Holm-corrected) ===\n")
pairs(em_color_int, adjust = "holm")

# ── 2. Intensity simple effects within each color ─────────────────────────────
# Which intensity levels differ from each other, separately for red and white?
em_int_color <- emmeans(simple_model1, ~ intensity | color, type = "response")

cat("\n=== Intensity pairwise contrasts within Red (Holm-corrected) ===\n")
pairs(em_int_color, adjust = "holm")

# ── 3. Distance pairwise contrasts ────────────────────────────────────────────
em_dist <- emmeans(simple_model1, ~ dist_cat, type = "response")

cat("\n=== Estimated marginal means: distance categories ===\n")
em_dist

cat("\n=== Distance pairwise contrasts (Holm-corrected) ===\n")
pairs(em_dist, adjust = "holm")
```

## Species Models

```{r species_models, warning=FALSE, message=FALSE}
source("R/11_species_models.R")
sp_results <- run_species_models(data_env, bat_theme, col_red, col_white)
```

```{r species_figs, warning=FALSE, message=FALSE}
sp_results$fig_s1   # Color effect by species
sp_results$fig_s2   # Intensity slopes by species
sp_results$fig_s3   # White:red ratio heatmap
```

## Figures

```{r figures, message=FALSE, warning=FALSE}
source("R/08_figures.R")

fig1; fig2; fig3; fig4; fig5; fig6; fig7; fig8
```

## Species Key

| Code | Species                                          |
|------|--------------------------------------------------|
| Myyu | Yuma myotis (possibly Myvo)                      |
| Myvo | *Myotis volans* \*                               |
| Myth | *Myotis thysanodes* (possibly Myev)              |
| Mylu | *Myotis lucifugus* \*                            |
| Myev | *Myotis evotis* \*                               |
| Myci | *Myotis ciliolabrum*                             |
| Myca | *Myotis californicus*                            |
| Lano | *Euderma maculatum*                              |
| Laci | *Lasiurus cinereus* \*                           |
| Euma | *Euderma maculatum*                              |
| Epfu | *Eptesicus fuscus* \*                            |
| Anpa | *Antrozous pallidus* (possibly *L. noctivagans*) |

------------------------------------------------------------------------

*At the end of every day:*

```         
git add .
git commit -m "updated detection pipeline"
git push
```
